### 2024-08-19 article_4.tex
article file uploaded

\documentclass{article}

\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\begin{document}

\title{Exponential Integrators for PDEs arising in meteorology} 
\maketitle
\author{Harvey Sutton}

\tableofcontents
\newpage

\section {Introduction}
%Statement of ODE
It is common within applied mathematics to encounter ODEs of the form
\begin{align}
\dot u(t) = Au(t) + N(u) \label{ODE}
\end{align}
Where $A$ is a large sparse matrix and $N$ denotes some nonlinear term.
As a result, finding accurate approximate solutions to these equations is of major importance.\\

%Method of lines
These ODEs also occure when solving PDEs numerically using the method of lines.
This arrives from discretising the PDE in space yielding an ODE with respect to time, with the form shown in \eqref{ODE}.\\
%ODE Solvers
Soving these ODEs numerically works by progressively stepping through time such as with the explicit forward Euler method, which is given as:
\begin{align*}
u(t+h) &= u(t) + h\dot u(t) = u(t) + h(Au(t) + N(u(t)))
\end{align*}
for a given step size $h$.
%Exponential Integrators
Another method (and the one we will be focusing on here) employs the use of the exponential integrator, which writes the solution to problem \eqref{ODE} over a given time step as:
\begin{align*}
u(t+h) &= e^{Ah}u(t) + \int_0^h e^{A(h-z)}N(u(t+z)) dz \label{ODE Solution}\\
\text{where } e^{A} &= \sum^{\inf}_{i = 0}\frac{A^n}{n!}
\end{align*}
%Computing matrix exponential for A sparse
An important requirement when using this method is to compute approximations of $e^{A}v$ efficiently and accurately for the given sparse matrix $A$ and vector $v$.
One approach would be first to compute $e^A$ and apply the resulting matrix to the vector $v$.
However, even for sparse $A$, $e^{A}$ will likely be dense. For large $A$ this will result in very high memory usage, which may not be practical when approximating solutions to \eqref{ODE Solution}.
One solution to reduce memory usage is to compute $e^{A}v$ directly without computing $e^{A}$ as an intermediary step.\\
Various methods have been proposed for this such as by Al-Mohy, Awad H. and Higham, Nicholas J\cite{AlMohy2011} which is currently used in the Scipy package as \verb|scipy.sparse.linalg.expm_multiply|.
Another class of methods known as Krylov subspace methods have gained traction recently\cite{Moler2003}.
These work by projecting the problem onto a space with fewer dimensions (the kyrlov subspace), computing the matrix exponential and then transforming back to the original space.
As the matrix exponential will be computed in the lower dimension kyrlov subspace the computation cost of this step will not be significant.

\newpage
\bibliographystyle{plain}
\bibliography{References.bib}

\end{document}

### 2024-08-20 steppers.py
add a starting point for implementing a 'stepper' in dune The SI example should be not too hard to modify to an exponential integrator

@@ -1,0 +1,293 @@
+# %% [markdown]
+#
+# # Exponential Integrators
+#
+# ## Setup
+
+# %%
+import matplotlib.pyplot as plt
+import numpy as np
+from scipy.sparse.linalg import cg, LinearOperator
+from scipy.sparse import identity, diags
+
+from dune.grid import structuredGrid as leafGridView
+from dune.fem.space import lagrange
+from dune.fem import integrate
+from dune.fem.operator import galerkin
+from ufl import ( TestFunction, TrialFunction, SpatialCoordinate, FacetNormal,
+                  dx, ds, div, grad, dot, inner, conditional, as_vector,
+                  exp, sin, pi )
+
+# %% [markdown]
+#
+# Simple utility function to show the result of a simulation
+
+# %%
+def printResult(method,error):
+    print(method,"L^2, H^1 error:",'{:0.5e}, {:0.5e}'.format(
+                 *[ np.sqrt(e) for e in integrate([error**2,inner(grad(error),grad(error))]) ]),
+          flush=True)
+
+# %% [markdown]
+# ## A base class for the steppers
+#
+# Assume the non-linear operator is given by `N` so we are solving
+# $u_t + N[u] = 0$
+
+# %%
+class BaseSolver:
+    def __init__(self, N, method="explicit"):
+        self.N = N
+        self.method = method
+        self.spc = N.domainSpace
+        self.un = self.spc.zero.copy()  # previous time step
+        self.res = self.un.copy()       # residual
+        self.shape = (self.spc.size,self.spc.size)
+
+        # inverse lumped mass matrix (see the explanation given in 'wave.py' tutorial example)
+        M = galerkin(dot(u,v)*dx).linear().as_numpy
+        Mdiag = M.sum(axis=1) # sum up the entries onto the diagonal
+        self.Minv = diags( 1/Mdiag.A1, shape=self.shape )
+        self.countN = 0
+        self.linIter = 0
+        # time step to use
+        self.tau = None
+
+        if not "expl" in self.method:     # for all non-explicit method we need the linearization
+            self.A = self.N.linear()
+            self.d = self.un.copy()       # search direction for non-linear method
+            self.I = identity(self.shape[0])
[…]

### 2024-08-20 steppers.py
[minor]

@@ -35,7 +35,7 @@ def printResult(method,error):
 # $u_t + N[u] = 0$
 
 # %%
-class BaseSolver:
+class BaseStepper:
     def __init__(self, N, method="explicit"):
         self.N = N
         self.method = method
@@ -129,9 +129,9 @@ class BaseSolver:
 
 # %%
 # solve d_t u + N(u) = 0 using u^{n+1} = u^n - tau N(u^n)
-class FESolver(BaseSolver):
+class FEStepper(BaseStepper):
     def __init__(self, N):
-        BaseSolver.__init__(self,N,"explicit")
+        BaseStepper.__init__(self,N,"explicit")
         self.name = "FE"
 
     def __call__(self, target, tau):
@@ -145,9 +145,9 @@ class FESolver(BaseSolver):
 
 # %%
 # solve d_t u + N(u) = 0 using u^{n+1} = u^n - tau N(u^{n+1})
-class BESolver(BaseSolver):
+class BEStepper(BaseStepper):
     def __init__(self, N, method):
-        BaseSolver.__init__(self,N,method)
+        BaseStepper.__init__(self,N,method)
         self.name = "BE"
 
     def __call__(self, target, tau):
@@ -189,9 +189,9 @@ class BESolver(BaseSolver):
 # So in each step we need to solve
 # ( I + tau A^n ) u^{n+1} = u^n - tau R^n(u^n)
 #                         = ( I + tau A^n) u^n - tau N(u^n)
-class SISolver(BaseSolver):
+class SIStepper(BaseStepper):
     def __init__(self, N):
-        BaseSolver.__init__(self,N, "quasiNewton")  # quasiNewton computes D in 'setup'
+        BaseStepper.__init__(self,N, "quasiNewton")  # quasiNewton computes D in 'setup'
         self.name = "SI1"
 
     def __call__(self, target, tau):
@@ -253,11 +253,11 @@ op = galerkin(a-bf-bg, domainSpace=space, rangeSpace=space)
 tauFE = 7e-5 # time step (FE fails with tau=8e-5 on the [80,80] grid)
 
 if False: # use FE
-    stepper = FESolver(op)
+    stepper = FEStepper(op)
     tau = tauFE
 else:
-    # stepper = BESolver(op, method="Newton")
-    stepper = SISolver(op)
+    # stepper = BEStepper(op, method="Newton")
+    stepper = SIStepper(op)
     tau = tauFE*10000
 
 # initial condition

### 2024-08-22 HeatEquations.py
Created Heat Equation

@@ -1,0 +1,42 @@
+import matplotlib.pyplot as plt
+import numpy as np
+from scipy.sparse.linalg import cg, LinearOperator
+from scipy.sparse import identity, diags
+
+from dune.grid import structuredGrid as leafGridView
+from dune.fem.space import lagrange, composite
+from dune.fem import integrate
+from dune.fem.operator import galerkin, molGalerkin
+from ufl import ( TestFunction, TrialFunction, SpatialCoordinate, FacetNormal,
+                  dx, ds, div, grad, dot, inner, exp, sin, conditional,
+                  sin, pi, as_vector)
+
+grid = leafGridView([0, 0], [1, 1], [100, 100])
+
+epsilon = 1
+
+
+space = lagrange(grid, order = 2)
+x = SpatialCoordinate(space)
+initial = sin(pi*x[0]) * sin(pi*x[1])
+u_h = space.interpolate(0, name = "u_h")
+u_h.interpolate(initial)
+u_h.plot()
+u_h_n = u_h.copy()
+
+u = TrialFunction(space)
+v = TestFunction(space)
+
+a = epsilon**2*dot(grad(u), grad(v)) * dx
+
+op = galerkin(a, rangeSpace=space, domainSpace=space)
+
+dt = 0.1
+T = 10000
+t = 0
+while t < T:
+    u_h_n.assign(u_h)
+    op(u_h, u_h_n)
+    u_h.as_numpy[:] -= dt * u_h_n.as_numpy[:]
+    t += dt
+u_h.plot()

### 2024-08-22 plot_data.py
Produced Data

@@ -9,7 +9,7 @@ e_threshold = 1e-10
 # %%
 with open("Experiment_Data.pickle", "rb") as file:
     data = pickle.load(file)
-ns = [1000, 10000, 100000]
+ns = [1000, 10000, 100000,1000000]
 # %%
 for n in ns:
     ndata = data[data["N"] == n]

### 2024-08-22 produce_data.py

@@ -26,8 +26,8 @@ def GetA(n):
 methods = {
     "Scipy": lambda A, v, m: expm_multiply(A,v),
     "Arnoldi": ArnoldiExp,
-    "Lanzcos": LanzcosExp,
-    "NBLA": NBLAExp
+    "Lanzcos": LanzcosExp#,
+    #"NBLA": NBLAExp
 }
 m = [1,2,3,4,5,6,7,8,9,10,20,30,40,50,60,70,80,90,100]
 n = [10,20,30,40,50,60,70,80,90,
@@ -90,7 +90,7 @@ for index, row in data.iterrows():
         # Gets an average
         total_time = 0
         total_error = 0
-        count = 1
+        count = 10
         for i in range(count):
             start = time.time()
             result = method(A, V, M)

### 2024-08-30 steppers.py
Added Exponential Integrator Stepper

@@ -18,6 +18,7 @@ from ufl import ( TestFunction, TrialFunction, SpatialCoordinate, FacetNormal,
                   dx, ds, div, grad, dot, inner, conditional, as_vector,
                   exp, sin, pi )
 
+from scipy.sparse.linalg import expm_multiply
 # %% [markdown]
 #
 # Simple utility function to show the result of a simulation
@@ -212,6 +213,38 @@ class SIStepper(BaseStepper):
         sol_coeff[:],_ = cg(self.D, res_coeff, callback=lambda x: self.callback(x) ) # note that this assumes a symmetric D
         return {"iterations":0, "linIter":self.linIter}
 
+# %%
+# solve d_t u + N(u) = 0 using u^{n+1} = e^{-tau A^n}u^n - tau R^n(u^n)
+# with A^n = DN(u^n) and R^n(u^n) = N(u^n) - A^n u^n, i.e., rewritting
+# N(u) = DN(u)u + (N(u) - DN(u)u)
+# So in each step we need to solve
+# u^{n+1} = e^{-tau A^n}u^n - tau R^n(u^n)
+class ExponentialStepper(BaseStepper):
+    def __init__(self, N, method, method_name = ""):
+        BaseStepper.__init__(self,N, "explicit")  # quasiNewton computes D in 'setup'
+        self.name = "ExpInt{0}".format(method_name)
+        self.A = self.N.linear()
+        self.exp_v = method
+
+    def __call__(self, target, tau):
+        
+        self.setup(target, tau)
+        # get numpy vectors for target, residual, search direction
+        sol_coeff = target.as_numpy
+        res_coeff = self.res.as_numpy
+        self.linIter = 0
+
+        # Compute 
+        self.evalN(target,self.res)
+        # Compute e^{tau A^n}u^n
+        sol_coeff[:] = self.exp_v(-self.Minv @ self.A.as_numpy * tau, sol_coeff[:])
+        # Add R^n(u^n) = N(u^n) - A^n u^n
+        sol_coeff -= res_coeff[:] - tau * self.Minv @ self.A.as_numpy @ self.un.as_numpy[:]
+
+        
+        return {"iterations":0, "linIter":self.linIter}
+
+
 # %% [markdown]
 #
 # # Numerical experiment
@@ -252,8 +285,9 @@ op = galerkin(a-bf-bg, domainSpace=space, rangeSpace=space)
 # %%
[…]

### 2024-08-30 article_4.tex

@@
 \usepackage{amsmath}
+\usepackage{amssymb}
+\usepackage{algorithm}
+\usepackage[noend]{algpseudocode}
 \begin{document}
@@
 As the matrix exponential will be computed in the lower dimension kyrlov subspace the computation cost of this step will not be significant.
+\section{Krylov Subspace Methods}
+In this section, we will look into existing krylov subspace methods and then compare how these methods perform for approximating $e^{A}v$.
 
+\subsection{Alogrithms}
+Here we give the algorithms for the 2 krylov subspace methods we will be comparing.
+Each of these algorithms takes a matrix $A\in \mathbb{R}^{n\times n}$, a vector $v \in \mathbb{R}^n$ and an integer $m$ that determines the number of dimensions of the krylov subspace used.
+From these algorithms we will get a matrix $H \in \mathbb{R}^{m\times m}$ and another matrix $V \in \mathbb{R}^{n\times m}$ such that $A \approx VHV^T$ and $VV^T = I$.
+From here we can get $Av \approx VHV^Tv = VH||v||e_1$.
+
+\begin{algorithm}
+\caption{Arnoldi \cite{Fan2018}} %find better citation
+\begin{algorithmic}
+\Procedure{Arnoldi}{$A, \hat v_1,m$}
+\State $v_0 = 0$
+\For{$j = 1,2,...,m$}	
+\For{$i = 1,2,...,j$}
+\State$h_{ij} \gets v_i^T A v_i$
+\EndFor
+\State$\theta_j \gets Av_j - \sum^j_{i=1} h_{ij}v_i$
+\State$h_{j+1,j} \gets ||\theta_j||$
+\State$v_{j+1} \gets \theta_j/h_{j+1,j}$
+\EndFor
+\EndProcedure
+\end{algorithmic}
+\end{algorithm}
+Here V is given by $v_1,...,v_m$ and H is give by $h_1,...,h_m$
+\begin{algorithm}
+\caption{Lanczos \cite{OJALVO1970}}
+\begin{algorithmic}
+\Procedure{Lanczos}{$A, \hat v_1,m$}
+\State $v_0 = 0$
+\For{$i \in \{1,2,...,m\}$}	
+\State$\beta_i \gets || \hat v_i ||$
+\State$v_i \gets \hat v_i / || \hat v_i ||$
+\State$\alpha_i = v_i^T A v_i$
+\State$\hat v_{i+1} = Av_i - \alpha_iv_i - \beta_iv_{i-1}$
+\EndFor
+\EndProcedure
+\end{algorithmic}
+\end{algorithm}\\
+Here $V$ is given by ${v_1,...,v_m}$ and $H$ is tridiagonal with the leading diagonal being $\alpha_1, ..., \alpha_m$ and the upper and lower diagonals being $\beta_2,...,\beta_m$.
 \newpage

### 2024-08-30 steppers.py
more work on exponential forward euler stepper

@@ -34,6 +34,7 @@ def printResult(method,error):
 #
 # Assume the non-linear operator is given by `N` so we are solving
 # $u_t + N[u] = 0$
+# 
 
 # %%
 class BaseStepper:
@@ -46,6 +47,9 @@ class BaseStepper:
         self.shape = (self.spc.size,self.spc.size)
 
         # inverse lumped mass matrix (see the explanation given in 'wave.py' tutorial example)
+        # u^n v dx = sum_j u^n_j phi_j phi_i dx = M u^n
+        # Mu^{n+1} = Mu^n - dt N(u^n) (for FE)
+        # u^{n+1} = u^n - dt M^{-1}N(u^n)
         M = galerkin(dot(u,v)*dx).linear().as_numpy
         Mdiag = M.sum(axis=1) # sum up the entries onto the diagonal
         self.Minv = diags( 1/Mdiag.A1, shape=self.shape )
@@ -84,6 +88,8 @@ class BaseStepper:
             self.D = self.DNapprox(self, u)
 
     # compute matrix D = tau M^{-1}DN[u] + I
+    # e.g. BE: u^{n+1} + tau M^{-1}N[u^{n+1}] = u^n
+    # lineaaized around u^n_k: (I + tau M^{-1}DN[u^n_k])u = u^n
     def matrix(self,u):
         self.N.jacobian(u,self.A)
         D = self.tau * self.Minv @ self.A.as_numpy
@@ -190,57 +196,65 @@ class BEStepper(BaseStepper):
 # So in each step we need to solve
 # ( I + tau A^n ) u^{n+1} = u^n - tau R^n(u^n)
 #                         = ( I + tau A^n) u^n - tau N(u^n)
+# u^{n+1} = u^n - tau ( I + tau A^n )^{-1} N(u^n)
 class SIStepper(BaseStepper):
     def __init__(self, N):
         BaseStepper.__init__(self,N, "quasiNewton")  # quasiNewton computes D in 'setup'
         self.name = "SI1"
 
-    def __call__(self, target, tau):
+    def explStep(self, target, tau):
         # this sets D = I + tau A^n and initialized u^n
         self.setup(target, tau)
         # get numpy vectors for target, residual, search direction
         sol_coeff = target.as_numpy
         res_coeff = self.res.as_numpy
-        d_coeff   = self.d.as_numpy
[…]

### 2024-08-30 steppers.py
switch to linear

@@ -288,10 +288,10 @@ exact = as_vector([ ( 1/2*(x[0]**2+x[1]**2) - 1/3*(x[0]**3 - x[1]**3) + 1 ) *
 
 # f = - laplace u      and grad(u).n = grad(exact).n on the boundary and f = - laplace(exact)
 # fv dx = - laplace u v dx = grad(u).grad(v) dx - grad(u).n v ds
-# fv dx + grad(exact).n v ds = grad(u).grad(v) dx
+# grad(u).grad(v) dx - (-laplace exact)v dx - grad(exact).n v ds = 0
 
-a  = ( inner(grad(u),grad(v)) + (1+dot(u,u))*dot(u,v) ) * dx
-bf = (-div(grad(exact[0])) + (1+exact[0]**2)*exact[0]) * v[0] * dx
+a  = ( inner(grad(u),grad(v)) ) * dx # + (1+dot(u,u))*dot(u,v) ) * dx
+bf = (-div(grad(exact[0])) ) * v[0] * dx # + (1+exact[0]**2)*exact[0]) * v[0] * dx
 bg = dot(grad(exact[0]),n)*v[0]*ds
 op = galerkin(a-bf-bg)
 
@@ -307,13 +307,13 @@ if False: # use FE
     # stepper = FEStepper(op)
     tau = tauFE
 else:
-    # stepper = BEStepper(op, method="Newton")
+    stepper = BEStepper(op, method="Newton")
     # stepper = SIStepper(op)
-    stepper = ExponentialStepper(op, method=expm_multiply, method_name="Scipy")
+    # stepper = ExponentialStepper(op, method=expm_multiply, method_name="Scipy")
     tau = tauFE*100 # *100 # *10000
 
 # initial condition
-u_h.interpolate( 1.5 )
+u_h.interpolate( exact )
 upd = u_h.copy()
 
 # time loop
@@ -332,6 +332,7 @@ while True:
     # we expect u^n -> exact for n->infty, i.e., u' -> 0
     # so we stop if upd = u^n - u^{n+1} is small
     upd.as_numpy[:] -= u_h.as_numpy[:]
+    upd.plot()
     update = np.sqrt( np.dot(upd.as_numpy,upd.as_numpy) )
     print(f"time step {n}, time {time}, N {stepper.countN}, iterations {info}, update {update}")
     if update < check:

### 2024-08-31 article_4.tex
Added Kyrlov Subspace Results

@@
 \usepackage[noend]{algpseudocode}
+\usepackage{graphicx}
+\usepackage{afterpage}
 \begin{document}
@@
 \end{algorithm}
-Here V is given by $v_1,...,v_m$ and H is give by $h_1,...,h_m$
+Here V is given by $v_1,...,v_m$ and H is give by $h_1,...,h_m$.\\
 \begin{algorithm}
@@
 \begin{algorithmic}
-\Procedure{Lanczos}{$A, \hat v_1,m$}
+\Procedure{Lanczos}{$A$ symetric$, \hat v_1,m$}
 \State $v_0 = 0$
@@
 Here $V$ is given by ${v_1,...,v_m}$ and $H$ is tridiagonal with the leading diagonal being $\alpha_1, ..., \alpha_m$ and the upper and lower diagonals being $\beta_2,...,\beta_m$.
+
+\subsection{Relevance To Matrix Exponentials}
+By using $A \approx VHV^T$ and the fact that $V^TV = I$ we can apply this to $e^A$ as follows:
+\begin{align*}
+e^A &= \sum^{\infty}_{i=0}\frac{A^i}{i\!}\\
+&= \sum^{\infty}_{i=0}\frac{(VHV^T)^i}{i\!} \\
+&= \sum^{\infty}_{i=0}\frac{VH^iV^T}{i\!} \\
+&\text {and then when computing $e^Av$ we get}\\
+e^Av &= (\sum^{\infty}_{i=0}\frac{VH^iV^T}{i\!})v \\
+&= \sum^{\infty}_{i=0}\frac{VH^iV^Tv}{i\!} \\
+&= \sum^{\infty}_{i=0}\frac{VH^ie_1}{i\!} \\
+&= V(\sum^{\infty}_{i=0}\frac{H^i}{i\!})e_1 \\
+&= Ve^He_1
+\end{align*}
+
+\subsection{Numerical Results}
+Here we run numerical experiments, computing $e^Av$ for a tridiagonal matrix $A$ of sizes $100,000$ and $1,000,000$.
+We will compare the methods with respect to computation time, error and size of the Kyrlov subspace.
+
+\begin{figure}
+
+    \centering
+    \begin{minipage}{0.5\textwidth}
+       \centering
+	  \includegraphics[width=\linewidth]{Plots/M v E Results for N=1000000.png}
+	  \label{fig:MEe7}
+       \centering
+	  \includegraphics[width=\linewidth]{Plots/M v Comp Time Results for N=1000000.png}
+	  \label{fig:MEe7}
+       \centering
+	  \includegraphics[width=\linewidth]{Plots/Comp Time v E Results for N=1000000.png}
+	  \label{fig:MEe7}
+    \end{minipage}\hfill
+    \begin{minipage}{0.5\textwidth}
+       \centering
+	  \includegraphics[width=\linewidth]{Plots/M v E Results for N=100000.png}
+	  \label{fig:MEe6}
+       \centering
+	  \includegraphics[width=\linewidth]{Plots/M v Comp Time Results for N=100000.png}
+	  \label{fig:MEe6}
+       \centering
+	  \includegraphics[width=\linewidth]{Plots/Comp Time v E Results for N=100000.png}
+	  \label{fig:MEe6}
+    \end{minipage}
+These results show that the Lanzcos algorithm outperforms the Arnoldi algorithm in both convergence rate and computational time.
+\end{figure}
+
+\begin{figure}
+Below we compare the time required for the different algorithms to get below an error of $10^{10}$ when computing $e^Av$ for various matrix sizes.
+\includegraphics[width=\linewidth]{Plots/time to get below an error of 1e-10.png}
+\clearpage
+Again we continue to see that the Lanzcos algorithm outperforms the Arnodli algorithm.
+\end{figure}
+\clearpage
 \newpage

### 2024-09-08 NBLA.py
added some example pde problems

@@ -31,6 +31,7 @@ def NBLA(A, B, C, m):
         delta = delta.T
 
         Sigma = W[:,j+1].T@V[:,j+1]
+        assert Sigma >= 0
         delta *= np.sqrt(Sigma)
         beta *= np.sqrt(Sigma) 
 
@@ -67,4 +68,3 @@ if __name__ == "__main__":
 
     print(NBLAExp(A, v, m))
     print(expm_multiply(A, v))
-

### 2024-09-08 allenCahn.py

@@ -1,0 +1,38 @@
+import numpy as np
+from ufl import *
+from dune.ufl import Space, Constant
+from dune.fem.function import gridFunction
+
+dimR = 1
+time = Constant(0,"time")
+sourceTime = Constant(0,"sourceTime")
+domain = [0, 0], [1, 1], [120, 120]
+
+space = Space(2,dimRange=dimR)
+x,u,v,n = ( SpatialCoordinate(space), TrialFunction(space), TestFunction(space), FacetNormal(space) )
+
+eps = Constant(0.01,"eps")
+theta = Constant(0.8,"theta")
+theta_c = Constant(1.6,"theta_c")
+
+a = ( eps**2 * inner(grad(u),grad(v))
+      - theta/2 * ln( (1-u[0])/(1+u[0]) ) * v[0]
+      - theta_c * dot(u,v)
+    ) * dx
+
+a = ( eps**2 * inner(grad(u),grad(v))
+      + (dot(u,u)-1) * dot(u,v)
+    ) * dx
+
+def test1(gridView):
+    h = 0.01
+    g0 = lambda x,x0,T: conditional(x-x0<-T/2,0,conditional(x-x0>T/2,0,sin(2*pi/T*(x-x0))**3))
+    G  = lambda x,y,x0,y0,T: g0(x,x0,T)*g0(y,y0,T)
+    return a, 20, 0.9*G(x[0],x[1],0.5,0.5,50*h), None
+
+def test2(gridView):
+    @gridFunction(gridView,name="random",order=1)
+    def u0(x):
+        return 1.8*np.random.rand(1)[0]-0.9
+    np.random.seed(100)
+    return a, 20, u0, None

### 2024-09-08 kiops.py

@@ -1,0 +1,341 @@
+# taken from https://github.com/uofs-simlab/pythOS/tree/main
+
+import math
+
+import numpy
+import scipy.linalg
+import numpy as np
+from scipy.sparse import diags, csc_matrix, csr_matrix, lil_matrix, lil_array
+import scipy.sparse
+from scipy.sparse.linalg import expm, expm_multiply
+from numpy.linalg import norm
+
+def kiops(τ_out, A, u, tol = 1e-10, m_init = 10, mmin = 10, mmax = 128, iop = 2, task1 = False):
+   """
+      kiops(tstops, A, u; kwargs...) -> (w, stats)
+
+   Evaluate a linear combinaton of the ``φ`` functions evaluated at ``tA`` acting on
+   vectors from ``u``, that is
+
+   ```math
+   w(i) = φ_0(t[i] A) u[0, :] + φ_1(t[i] A) u[1, :] + φ_2(t[i] A) u[2, :] + ...
+   ```
+
+   The size of the Krylov subspace is changed dynamically during the integration.
+   The Krylov subspace is computed using the incomplete orthogonalization method.
+
+   Arguments:
+   - `τ_out`    - Array of `τ_out`
+   - `A`        - the matrix argument of the ``φ`` functions
+   - `u`        - the matrix with rows representing the vectors to be multiplied by the ``φ`` functions
+
+   Optional arguments:
+   - `tol`      - the convergence tolerance required (default: 1e-7)
+   - `mmin`, `mmax` - let the Krylov size vary between mmin and mmax (default: 10, 128)
+   - `m`        - an estimate of the appropriate Krylov size (default: mmin)
+   - `iop`      - length of incomplete orthogonalization procedure (default: 2)
+   - `task1`     - if true, divide the result by 1/T**p
+
+   Returns:
+   - `w`      - the linear combination of the ``φ`` functions evaluated at ``tA`` acting on the vectors from ``u``
+   - `stats[0]` - number of substeps
+   - `stats[1]` - number of rejected steps
+   - `stats[2]` - number of Krylov steps
+   - `stats[3]` - number of matrix exponentials
+   - `stats[4]` - Error estimate
+   - `stats[5]` - the Krylov size of the last substep
+
+   `n` is the size of the original problem
[…]

### 2024-09-08 model_euler_atmos.py

@@ -1,0 +1,212 @@
+import numpy
+
+from dune.common import FieldVector
+from dune.grid import reader, cartesianDomain
+from dune.fem.function import uflFunction
+from ufl import *
+from dune.ufl import cell, Constant
+
+
+def model(dim=2, **kwargs):
+    problem = ""
+    cmpE = dim+1
+    class Model:
+        dimRange = dim+2
+        cmpE = dim+1
+
+        atmos = None
+        g   = 9.80665 # gravity force [m/ss]
+        T0  = 250.    # surface temperature [K]
+        p0  = 100000. # surface pressure [Pa]
+        c_p = 1005.   # todo: docme
+        c_v = 717.95  # todo: docme
+        N   = 0.01    # Brunt-Vaisala frequency (>0: stratified, 0: neutral, <0: isothermal atmosphere)
+        mu  = Constant(0., "kinetic_viscosity")  # kinetic viscosity [mm/s]
+        Re  = 1.      # Reynold's number
+        Pr  = 1.      # Prandtl's number
+
+        c_pd = 1005.  # specific heat of dry air at constant pressure, J/(kgK)=mm/(Kss)
+        c_vd = 717.95 # specific heat of dry air at constant volume
+
+        R_d = (c_pd - c_vd)
+        kappa = R_d / c_pd
+        gamma = (c_pd / c_vd )
+
+        hSpeed0 = 0 # 20.
+
+        # helper vars
+        _gNcp = g*g/(N*N*c_p)
+        _NNg  = N*N/g
+
+        def bgPrim(t,x):
+            # height in atmosphere
+            z = x[dim-1]
+            NNgz = Model._NNg * z
+
+            ## N > 0, TODO, other cases
+            # assert Model.N > 0, "Only N>0 implemented"
+            assert Model.atmos is not None
+
+            if Model.atmos == "stable":
+                # assert Model.N > 0, "N does not fit to selected atmosphere"
+                ## is this the 'stable atmosphere' from metstroem project?
+                T = Model._gNcp + (Model.T0 - Model._gNcp) * exp( NNgz )
+                p = Model.p0 * exp( 1./Model.kappa * (ln(T/Model.T0) - NNgz) )
+                theta = T * pow( p/Model.p0, -Model.kappa )
+            elif Model.atmos == "isothermal":
+                # where is Model.N used in this atmosphere?
+                # The derived quantities gNcp and NNgz are only used in
[…]

### 2024-09-08 parabolicTest.py

@@ -1,0 +1,39 @@
+from ufl import *
+from dune.ufl import Space, Constant
+
+dimR = 1
+time = Constant(0,"time")
+sourceTime = Constant(0,"sourceTime")
+domain = [0, 0], [1, 1], [80, 80]
+
+space = Space(2,dimRange=dimR)
+x,u,v,n = ( SpatialCoordinate(space), TrialFunction(space), TestFunction(space), FacetNormal(space) )
+
+def model(exact,dtExact,nonLinear):
+    a  = ( inner(grad(u),grad(v)) + dot(nonLinear(u),v) ) * dx
+    bf = lambda t: ( ( dtExact(t)[0] - div(grad(exact(t)[0])) ) * v[0]
+                     + dot(nonLinear(exact(t)),v) 
+                   ) * dx 
+    bg = lambda t: dot(grad(exact(t)[0]),n)*v[0]*ds
+    return a-bf(sourceTime)-bg(sourceTime)
+
+def paraTest0(gridView):
+    freq = 2*pi
+    xExact = as_vector([ cos(freq*x[0]) ])
+    exact = lambda t: exp(-freq**2*t) * xExact
+    dtExact = lambda t: -freq**2*exp(-freq**2*t) * xExact
+    return model(exact,dtExact, lambda u: as_vector([0])), 0.5, exact(time), exact
+
+def paraTest1(gridView):
+    freq = 6*pi
+    xExact = as_vector([ cos(freq*x[0]) ])
+    exact = lambda t: exp(-freq**2*t) * xExact
+    dtExact = lambda t: -freq**2*exp(-freq**2*t) * xExact
+    return model(exact,dtExact, lambda u: (1+dot(u,u))*u), 0.5, exact(0), exact
+
+def paraTest2(gridView):
+    freq = 2*pi
+    xExact = as_vector([ cos(freq*x[0]) ])
+    exact = lambda t: exp(-freq**2*t) * xExact
+    dtExact = lambda t: -freq**2*exp(-freq**2*t) * xExact
+    return model(exact,dtExact, lambda u: (1+dot(u,u))*u), 0.5, exact(time), exact

### 2024-09-08 risingbubble.py

@@ -1,0 +1,141 @@
+import sys
+
+from dune.fem.function import gridFunction
+from dune.ufl import cell, Constant
+from ufl import SpatialCoordinate, sqrt, exp, sin, pi, conditional, as_vector
+
+from model_euler_atmos import model as AtmosphericModel
+from model_bgfix import model as BgFixModel
+
+
+def RisingBubble(dim=2):
+    Model = AtmosphericModel(dim=dim)
+    Model.atmos = "neutral"
+    Model.hSpeed0 = 0.
+    Model.T0 = 303.15
+    #Model.K = 0.1
+
+    x = SpatialCoordinate(cell(dim))
+    Model.dx = Constant(0., name="dx")
+
+    z   = x[dim-1] # z-coordinate
+    #x_c = 533.     # x-center of perturbation ball
+    x_c = 500. + 0.5*Model.dx  # x-center of perturbation ball
+    z_c = 520.     # z-center of perturbation ball
+    a   = 50       # radius of perturbation ball
+
+    r = sqrt((x[0] - x_c)**2 + (z - z_c)**2)
+
+    BG, p, T, theta, v = Model.bg(0, x)
+
+    A = 0.5
+    s = 100.
+
+    deltaTheta = A * conditional(r <= a,
+        1.0,
+        exp( -(r-a)**2/s**2)
+    )
+    newTheta = theta + deltaTheta
+
+    # introduce pot. temp. to conservative variables keeping
+    # the pressure equal to the background pressure
+
+    rho = pow( Model.p0, Model.kappa ) * pow( p, 1./Model.gamma ) / Model.R_d / newTheta
+
+    Model.U0 = Model.toCons(as_vector([rho] + v + [newTheta]))
+
+    Model.boundary = {1: lambda t,x,U: Model.reflect(U,[-1, 0]),
+                      2: lambda t,x,U: Model.reflect(U,[ 1, 0]),
+                      3: lambda t,x,U: Model.reflect(U,[ 0,-1]),
+                      4: lambda t,x,U: Model.reflect(U,[ 0, 1])}
+
+
+    Model.domain = [0]*dim, [1000]*(dim-1)+[2000], [40]*(dim-1)+[80]
+    Model.endTime = 1200.
+    Model.name = "RisingBubble"
+
+    return BgFixModel(Model, dim)
+
+import sys
+import numpy as np
+import matplotlib.pyplot as plt
+from dune.grid import structuredGrid
+from dune.fem.space import dglagrange, finiteVolume
[…]

### 2024-09-08 steppers.py

@@ -17,16 +17,20 @@ from scipy.sparse.linalg import expm_multiply
 import random
 
 from Stable_Lanzcos import LanzcosExp
-expm_lanzcos = lambda A,x: LanzcosExp(A,x,5)
+expm_lanzcos = [lambda A,x: LanzcosExp(A,x,5),"Lanzcos"]
+from NBLA import NBLAExp
+expm_nbla = [lambda A,x: NBLAExp(A,x,5),"NBLA"]
+from Arnoldi import ArnoldiExp 
+expm_arnoldi = [lambda A,x: ArnoldiExp(A,x,5),"Arnoldi"]
+from kiops import KiopsExp
+expm_kiops = [lambda A,x: KiopsExp(A,x),"Kiops"]
 
 from dune.grid import structuredGrid as leafGridView
 from dune.fem.space import lagrange
 from dune.fem import integrate
 from dune.fem.operator import galerkin
-from dune.ufl import Constant
-from ufl import ( TestFunction, TrialFunction, SpatialCoordinate, FacetNormal,
-                  dx, ds, div, grad, dot, inner, conditional, as_vector,
-                  exp, sin, cos, pi )
+from dune.fem.function import gridFunction
+from ufl import dx, dot, inner, grad, TestFunction, TrialFunction
 
 # %% [markdown]
 #
@@ -48,7 +52,8 @@ def printResult(time,error,*args):
 # %%
 class BaseStepper:
     # method is 'explicit', 'approx', 'exact'
-    def __init__(self, N, method="explicit"):
+    # mass is 'lumped', 'exact', 'identitiy'
+    def __init__(self, N, *, method="explicit", mass="lumped"):
         self.N = N
         self.method = method
         self.spc = N.domainSpace
@@ -56,13 +61,24 @@ class BaseStepper:
         self.res = self.un.copy()       # residual
         self.shape = (self.spc.size,self.spc.size)
 
-        # inverse lumped mass matrix (see the explanation given in 'wave.py' tutorial example)
-        # u^n v dx = sum_j u^n_j phi_j phi_i dx = M u^n
-        # Mu^{n+1} = Mu^n - dt N(u^n) (for FE)
-        # u^{n+1} = u^n - dt M^{-1}N(u^n)
-        M = galerkin(dot(u,v)*dx).linear().as_numpy
-        Mdiag = M.sum(axis=1) # sum up the entries onto the diagonal
[…]

### 2024-09-09 allenCahn.py
[minor]

@@ -15,10 +15,12 @@ eps = Constant(0.01,"eps")
 theta = Constant(0.8,"theta")
 theta_c = Constant(1.6,"theta_c")
 
+"""
 a = ( eps**2 * inner(grad(u),grad(v))
       - theta/2 * ln( (1-u[0])/(1+u[0]) ) * v[0]
       - theta_c * dot(u,v)
     ) * dx
+"""
 
 a = ( eps**2 * inner(grad(u),grad(v))
       + (dot(u,u)-1) * dot(u,v)

### 2024-09-09 parabolicTest.py

@@ -18,7 +18,7 @@ def model(exact,dtExact,nonLinear):
     return a-bf(sourceTime)-bg(sourceTime)
 
 def paraTest0(gridView):
-    freq = 2*pi
+    freq = 6*pi
     xExact = as_vector([ cos(freq*x[0]) ])
     exact = lambda t: exp(-freq**2*t) * xExact
     dtExact = lambda t: -freq**2*exp(-freq**2*t) * xExact
@@ -32,8 +32,8 @@ def paraTest1(gridView):
     return model(exact,dtExact, lambda u: (1+dot(u,u))*u), 0.5, exact(0), exact
 
 def paraTest2(gridView):
-    freq = 2*pi
+    freq = 6*pi
     xExact = as_vector([ cos(freq*x[0]) ])
-    exact = lambda t: exp(-freq**2*t) * xExact
-    dtExact = lambda t: -freq**2*exp(-freq**2*t) * xExact
+    exact = lambda t: cos(freq**2*t) * xExact
+    dtExact = lambda t: -freq**2*sin(freq**2*t) * xExact
     return model(exact,dtExact, lambda u: (1+dot(u,u))*u), 0.5, exact(time), exact

### 2024-09-09 allenCahn.py
[minor]

@@ -6,7 +6,7 @@ from dune.fem.function import gridFunction
 dimR = 1
 time = Constant(0,"time")
 sourceTime = Constant(0,"sourceTime")
-domain = [0, 0], [1, 1], [120, 120]
+domain = [0, 0], [1, 1], [60, 60]
 
 space = Space(2,dimRange=dimR)
 x,u,v,n = ( SpatialCoordinate(space), TrialFunction(space), TestFunction(space), FacetNormal(space) )
@@ -26,15 +26,18 @@ a = ( eps**2 * inner(grad(u),grad(v))
       + (dot(u,u)-1) * dot(u,v)
     ) * dx
 
+"""
 def test1(gridView):
     h = 0.01
     g0 = lambda x,x0,T: conditional(x-x0<-T/2,0,conditional(x-x0>T/2,0,sin(2*pi/T*(x-x0))**3))
     G  = lambda x,y,x0,y0,T: g0(x,x0,T)*g0(y,y0,T)
     return a, 20, 0.9*G(x[0],x[1],0.5,0.5,50*h), None
+"""
 
 def test2(gridView):
+    tauFE = 0.8   # AC on level=0
     @gridFunction(gridView,name="random",order=1)
     def u0(x):
         return 1.8*np.random.rand(1)[0]-0.9
     np.random.seed(100)
-    return a, 20, u0, None
+    return a, 24, tauFE, u0, None

### 2024-09-09 parabolicTest.py

@@ -18,22 +18,25 @@ def model(exact,dtExact,nonLinear):
     return a-bf(sourceTime)-bg(sourceTime)
 
 def paraTest0(gridView):
+    tauFE = 7e-5 # check if this is correct
     freq = 6*pi
     xExact = as_vector([ cos(freq*x[0]) ])
     exact = lambda t: exp(-freq**2*t) * xExact
     dtExact = lambda t: -freq**2*exp(-freq**2*t) * xExact
-    return model(exact,dtExact, lambda u: as_vector([0])), 0.5, exact(time), exact
+    return model(exact,dtExact, lambda u: as_vector([0])), 0.5, tauFE, exact(time), exact
 
 def paraTest1(gridView):
+    tauFE = 7e-5 # check if this is correct
     freq = 6*pi
     xExact = as_vector([ cos(freq*x[0]) ])
     exact = lambda t: exp(-freq**2*t) * xExact
     dtExact = lambda t: -freq**2*exp(-freq**2*t) * xExact
-    return model(exact,dtExact, lambda u: (1+dot(u,u))*u), 0.5, exact(0), exact
+    return model(exact,dtExact, lambda u: (1+dot(u,u))*u), 0.5, tauFE, exact(0), exact
 
 def paraTest2(gridView):
+    tauFE = 7e-5 # check if this is correct
     freq = 6*pi
     xExact = as_vector([ cos(freq*x[0]) ])
     exact = lambda t: cos(freq**2*t) * xExact
     dtExact = lambda t: -freq**2*sin(freq**2*t) * xExact
-    return model(exact,dtExact, lambda u: (1+dot(u,u))*u), 0.5, exact(time), exact
+    return model(exact,dtExact, lambda u: (1+dot(u,u))*u), 0.5, tauFE, exact(time), exact

### 2024-09-09 risingbubble.py

@@ -63,19 +63,9 @@ from dune.grid import structuredGrid
 from dune.fem.space import dglagrange, finiteVolume
 from dune.femdg import femDGModels, femDGOperator, advectionNumericalFlux
 from dune.femdg.rk import femdgStepper
-from steppers import BEStepper, FEStepper, ExponentialStepper, SIStepper
-from steppers import expm_kiops, expm_lanzcos, expm_nbla, expm_arnoldi, expm_multiply
+from steppers import steppersDict
 
 if __name__ == "__main__":
-    steppersDict = {"FE": (FEStepper,{}),
-                    "BE": (BEStepper,{"ftol":1e-3,"verbose":False}),
-                    "SI": (SIStepper,{}),
-                    "EXPSCI": (ExponentialStepper, {"exp_v":expm_multiply}),
-                    "EXPLAN": (ExponentialStepper, {"exp_v":expm_lanzcos}),
-                    "EXPNBLA": (ExponentialStepper, {"exp_v":expm_nbla}),
-                    "EXPARN": (ExponentialStepper, {"exp_v":expm_arnoldi}),
-                    "EXPKIOPS": (ExponentialStepper, {"exp_v":expm_kiops}),
-                   }
     stepperFct, args = steppersDict[sys.argv[1]]
     if len(sys.argv) > 2:
         cfl = int(sys.argv[2])
@@ -108,34 +98,33 @@ if __name__ == "__main__":
     info = stepper(target=u_h, tau=1e-5)
     u_h.interpolate(Model.U0)
 
-    # stepper = femdgStepper(order=1, operator=op)
-
     # time loop
     # figure out a first tau
 
     t = 0
     n = 0
+    fileCount = 0
     plotTime = 0.5
     nextTime = 0.5
 
     u_h.plot(gridLines=None, block=False)
-    plt.savefig(outputName(n))
+    plt.savefig(outputName(fileCount))
+    fileCount += 1
 
     while t < 400:
         op.setTime(t)
         tau = op.localTimeStepEstimate[0]*cfl
-
-        """
-        t += stepper(u_h)
-        """
         info = stepper(target=u_h, tau=tau)
         t += tau
 
         assert not np.isnan(u_h.as_numpy).any()
         n += 1
         print(f"time step {n}, time {t}, tau {tau}, calls {op.info()}")
-        if t>plotTime:
+        if True: # t>plotTime:
             u_h.plot(gridLines=None, block=False)
-            plt.savefig(outputName(n))
[…]

### 2024-09-09 steppers.py

@@ -17,17 +17,19 @@ from scipy.sparse.linalg import expm_multiply
 import random
 
 from Stable_Lanzcos import LanzcosExp
-expm_lanzcos = [lambda A,x: LanzcosExp(A,x,5),"Lanzcos"]
+expm_lanzcos = [lambda A,x: LanzcosExp(A,x,m),"Lanzcos"]
 from NBLA import NBLAExp
-expm_nbla = [lambda A,x: NBLAExp(A,x,5),"NBLA"]
+expm_nbla = [lambda A,x,m: NBLAExp(A,x.m),"NBLA"]
 from Arnoldi import ArnoldiExp 
-expm_arnoldi = [lambda A,x: ArnoldiExp(A,x,5),"Arnoldi"]
+expm_arnoldi = [lambda A,x,m: ArnoldiExp(A,x,m),"Arnoldi"]
 from kiops import KiopsExp
-expm_kiops = [lambda A,x: KiopsExp(A,x),"Kiops"]
+expm_kiops = [lambda A,x,m: KiopsExp(A,x),"Kiops"]
 
-from dune.grid import structuredGrid as leafGridView
+from dune.grid import cartesianDomain
+from dune.alugrid import aluCubeGrid as leafGridView
 from dune.fem.space import lagrange
-from dune.fem import integrate
+from dune.fem import integrate, threading, globalRefine
+from dune.fem.view import adaptiveLeafGridView as view
 from dune.fem.operator import galerkin
 from dune.fem.function import gridFunction
 from ufl import dx, dot, inner, grad, TestFunction, TrialFunction
@@ -53,6 +55,10 @@ def printResult(time,error,*args):
 class BaseStepper:
     # method is 'explicit', 'approx', 'exact'
     # mass is 'lumped', 'exact', 'identitiy'
+    # Change:
+    # - define class that wraps an operator 'N' and returns object with same
+    #   interface but including M^{-1}
+    # - put N on right hand side
     def __init__(self, N, *, method="explicit", mass="lumped"):
         self.N = N
         self.method = method
@@ -209,7 +215,7 @@ class FEStepper(BaseStepper):
 # solve d_t u + N(u) = 0 using u^{n+1} = u^n - tau N(u^{n+1})
 class BEStepper(BaseStepper):
     def __init__(self, N, *, method="approx", mass='lumped',
-                 ftol=1e-6, verbose=True):
+                 ftol=1e-6, verbose=False):
[…]

### 2024-09-09 risingbubble.py
[minor]

@@ -67,7 +67,7 @@ from steppers import steppersDict
 
 if __name__ == "__main__":
     stepperFct, args = steppersDict[sys.argv[1]]
-    if len(sys.argv) > 2:
+    if len(sys.argv) >= 3:
         cfl = int(sys.argv[2])
         outputName = lambda n: f"risingbubble_{sys.argv[1]}{cfl}_{n}"
     else:
@@ -91,6 +91,11 @@ if __name__ == "__main__":
                       threading=True,
                       defaultQuadrature=True)
     op = femDGOperator(models, space, **operator_kwargs)
+
+    if len(sys.argv) >= 4:
+        if "exp_v" in args.keys():
+            m = int(sys.argv[3])
+            args["expv_args"] = {"m":m}
 
     stepper = stepperFct(op, mass='identity', **args)
 

### 2024-09-09 steppers.py

@@ -17,7 +17,7 @@ from scipy.sparse.linalg import expm_multiply
 import random
 
 from Stable_Lanzcos import LanzcosExp
-expm_lanzcos = [lambda A,x: LanzcosExp(A,x,m),"Lanzcos"]
+expm_lanzcos = [lambda A,x,m: LanzcosExp(A,x,m),"Lanzcos"]
 from NBLA import NBLAExp
 expm_nbla = [lambda A,x,m: NBLAExp(A,x.m),"NBLA"]
 from Arnoldi import ArnoldiExp 
@@ -299,18 +299,18 @@ class SIStepper(BaseStepper):
 #         = e^{-A^n tau} ( (I + tau A^n)u^n - tau N(u^n) )
 
 class ExponentialStepper(SIStepper):
-    def __init__(self, N, exp_v, *, exp_args, method='approx', mass='lumped'):
+    def __init__(self, N, exp_v, *, expv_args, method='approx', mass='lumped'):
         SIStepper.__init__(self,N, method=method, mass=mass)
         self.name = f"ExpInt({method},{exp_v[1]})"
         self.exp_v = exp_v[0]
-        self.exp_args = exp_args
+        self.expv_args = expv_args
 
     def __call__(self, target, tau):
         # u^* = (I + tau A^n)u^n - tau N(u^n)
         # Note: the call method on the base class calls 'setup' which computes the right A^n
         self.explStep(target,tau)
         # Compute e^{-tau A^n}u^*
-        target.as_numpy[:] = self.exp_v(- self.A, self.res.as_numpy, **self.exp_args)
+        target.as_numpy[:] = self.exp_v(- self.A, self.res.as_numpy, **self.expv_args)
         return {"iterations":0, "linIter":0}
 
 steppersDict = {"FE": (FEStepper,{}),
@@ -350,9 +350,9 @@ if __name__ == "__main__":
     # %%
     stepperFct, args = steppersDict[sys.argv[1]]
     if len(sys.argv) >= 5:
-        if exp_v in args.keys():
+        if "exp_v" in args.keys():
             m = int(sys.argv[4])
-            exp_v["m"] = m
+            args["expv_args"] = {"m":m}
 
     factor = float(sys.argv[2])
     tau = tauFE * factor

### 2024-09-10 plot_data.py
Changed values for n and stored data in csv format

@@ -7,9 +7,8 @@ import pickle as pickle
 e_threshold = 1e-10
 
 # %%
-with open("Experiment_Data.pickle", "rb") as file:
-    data = pickle.load(file)
-ns = [1000, 10000, 100000,1000000]
+data = pd.read_csv("Experiment_Data.csv")
+ns = data["N"].unique()
 # %%
 for n in ns:
     ndata = data[data["N"] == n]

### 2024-09-10 produce_data.py

@@ -30,13 +30,9 @@ methods = {
     #"NBLA": NBLAExp
 }
 m = [1,2,3,4,5,6,7,8,9,10,20,30,40,50,60,70,80,90,100]
-n = [10,20,30,40,50,60,70,80,90,
-     100,200,300,400,500,600,700,800,900,
-     1000,2000,3000,4000,5000,6000,7000,8000,9000,
-     10000,20000,30000,40000,50000,60000,70000,80000,90000,
-     100000,200000,300000,400000,500000,600000,700000,800000,900000,
-     1000000
-     ]
+n = [2**i for i in range(0, 20)]
+print(n)
+
 v = np.random.rand(np.max(n))
 method_names = list(methods.keys())
 mlen = len(m)
@@ -108,6 +104,5 @@ data["Computation Time"] = computation_times
 
 data.dropna()
 
-with open("Experiment_Data.pickle", "wb") as file:
-    pickle.dump(data, file, pickle.HIGHEST_PROTOCOL)
+data.to_csv("Experiment_Data.csv")
 

### 2024-09-10 plot_data.py
added scipy label

@@ -29,7 +29,7 @@ for n in ns:
         plt.figure(2)
         if method == "Scipy":
             mean_computation_time = np.mean(methoddata["Computation Time"])
-            plt.hlines(mean_computation_time, np.min(ndata["M"]), np.max(ndata["M"]), colors="r", linestyles="dashed")
+            plt.hlines(mean_computation_time, np.min(ndata["M"]), np.max(ndata["M"]), colors="r", linestyles="dashed", label="Scipy")
         else:
             plt.loglog(methoddata["M"], methoddata["Computation Time"], label= method)
         plt.xlabel("M")
@@ -39,7 +39,7 @@ for n in ns:
         
         plt.figure(3)
         if method == "Scipy":
-            plt.vlines(mean_computation_time, np.min(ndata["Error"]), np.max(ndata["Error"]), colors="r", linestyles="dashed")
+            plt.vlines(mean_computation_time, np.min(ndata["Error"]), np.max(ndata["Error"]), colors="r", linestyles="dashed", label="Scipy")
         else:
             plt.loglog(methoddata["Computation Time"], methoddata["Error"], label= method)
         plt.xlabel("Computation Time $s$")

### 2024-09-10 produce_data.py
A now scales with N squared

@@ -21,7 +21,7 @@ from NBLA import NBLAExp
 def GetA(n):
     A = diags([-1,2,-1], [-1,0,1], shape=(n,n))
     A = csc_matrix(A)
-    return A
+    return A * n**2
 
 methods = {
     "Scipy": lambda A, v, m: expm_multiply(A,v),

### 2024-09-10 plot_data.py
Proper labeling of axises

@@ -21,7 +21,7 @@ for n in ns:
             mean_Error = np.mean(methoddata["Error"])
         else:
             plt.loglog(methoddata["M"], methoddata["Error"], label= method)
-        plt.xlabel("M")
+        plt.xlabel("Dimention of Kyrlov subspace $m$")
         plt.ylabel("Error")
         plt.title("Graph Compairing M and Error for N={0}".format(n))
         plt.legend()
@@ -32,7 +32,7 @@ for n in ns:
             plt.hlines(mean_computation_time, np.min(ndata["M"]), np.max(ndata["M"]), colors="r", linestyles="dashed", label="Scipy")
         else:
             plt.loglog(methoddata["M"], methoddata["Computation Time"], label= method)
-        plt.xlabel("M")
+        plt.xlabel("Dimention of Kyrlov subspace $m$")
         plt.ylabel("Computation Time $s$")
         plt.title("Graph Compairing M and Computation Time for N={0}".format(n))
         plt.legend()

### 2024-09-10 plot_data.py
Fixed plot colours and scaling

@@ -60,13 +60,12 @@ for n in ns:
 # %% Plotting time for error to be bellow given bound
 cut_data = data[data["Error"] < e_threshold]
 ns = cut_data["N"].unique()
-for method in cut_data["Method"].unique():
+for method in data["Method"].unique():
     method_data = cut_data[cut_data["Method"] == method]
     result = []
     for n in ns:
         ndata = method_data[method_data["N"] == n]
         if method == "Scipy":
-            #result.append(ndata["Computation Time"].mean())
             continue
         try:
             result.append(ndata["Computation Time"].to_list()[ndata["Error"].argmin()])

### 2024-09-10 produce_data.py

@@ -9,6 +9,7 @@ import scipy.sparse
 from Arnoldi import ArnoldiExp
 from Stable_Lanzcos import LanzcosExp
 from NBLA import NBLAExp
+from kiops import KiopsExp
 
 # Collects data on performance of different methods
 # Independant Variables:
@@ -21,16 +22,17 @@ from NBLA import NBLAExp
 def GetA(n):
     A = diags([-1,2,-1], [-1,0,1], shape=(n,n))
     A = csc_matrix(A)
-    return A * n**2
+    return A / n**2
 
 methods = {
     "Scipy": lambda A, v, m: expm_multiply(A,v),
     "Arnoldi": ArnoldiExp,
     "Lanzcos": LanzcosExp#,
+    #"Kiops": lambda A, V, m: KiopsExp(A,V)
     #"NBLA": NBLAExp
 }
 m = [1,2,3,4,5,6,7,8,9,10,20,30,40,50,60,70,80,90,100]
-n = [2**i for i in range(0, 20)]
+n = [2**i for i in range(0, 8)]
 print(n)
 
 v = np.random.rand(np.max(n))
@@ -57,6 +59,7 @@ computation_times = []
 true_values = []
 scipy_computing_time =[]
 for N in n:
+    print("Computing {0}".format(N))
     A = GetA(N)
     V = v[0:N]
     start = time.time()

### 2024-09-11 steppers.py
allow to set 'm' in exp steppers, fix bug in scipy stepper, set some parameters in kiops

@@ -13,13 +13,14 @@ from scipy.linalg import norm
 from scipy.sparse.linalg import lgmres, gmres, LinearOperator, aslinearoperator
 from scipy.sparse import identity, diags
 from scipy.optimize import newton_krylov, KrylovJacobian
-from scipy.sparse.linalg import expm_multiply
 import random
 
+from scipy.sparse.linalg import expm_multiply
+expm_sci = [lambda A,x,m: expm_multiply(A,x),"Scipy"]
 from Stable_Lanzcos import LanzcosExp
 expm_lanzcos = [lambda A,x,m: LanzcosExp(A,x,m),"Lanzcos"]
 from NBLA import NBLAExp
-expm_nbla = [lambda A,x,m: NBLAExp(A,x.m),"NBLA"]
+expm_nbla = [lambda A,x,m: NBLAExp(A,x,m),"NBLA"]
 from Arnoldi import ArnoldiExp 
 expm_arnoldi = [lambda A,x,m: ArnoldiExp(A,x,m),"Arnoldi"]
 from kiops import KiopsExp
@@ -316,7 +317,7 @@ class ExponentialStepper(SIStepper):
 steppersDict = {"FE": (FEStepper,{}),
                 "BE": (BEStepper,{}),
                 "SI": (SIStepper,{}),
-                "EXPSCI": (ExponentialStepper, {"exp_v":expm_multiply}),
+                "EXPSCI": (ExponentialStepper, {"exp_v":expm_sci}),
                 "EXPLAN": (ExponentialStepper, {"exp_v":expm_lanzcos}),
                 "EXPNBLA": (ExponentialStepper, {"exp_v":expm_nbla}),
                 "EXPARN": (ExponentialStepper, {"exp_v":expm_arnoldi}),

### 2024-09-11 NBLA.py
Merge

@@ -31,6 +31,7 @@ def NBLA(A, B, C, m):
         delta = delta.T
 
         Sigma = W[:,j+1].T@V[:,j+1]
+        assert Sigma >= 0
         delta *= np.sqrt(Sigma)
         beta *= np.sqrt(Sigma) 
 
@@ -49,7 +50,9 @@ def NBLAExp(A, v_1, m):
     H, V, W, beta = NBLA(A, v_1, v_1, m)
     e_1 = np.zeros((m))
     e_1[0] = 1
-    return V @ expm_multiply(H, e_1) * beta
+    ret = V @ expm_multiply(H, e_1) * beta
+    # print(v_1.dot(v_1), ret.dot(ret))
+    return ret # V@expm_multiply(H,e_1)*np.linalg.norm(v_1)
 
 if __name__ == "__main__":
     np.set_printoptions(linewidth=200)
@@ -67,4 +70,3 @@ if __name__ == "__main__":
 
     print(NBLAExp(A, v, m))
     print(expm_multiply(A, v))
-

### 2024-09-11 Stable_Lanzcos.py

@@ -3,12 +3,12 @@ import numpy as np
 from numpy.linalg import norm
 import scipy as scipy
 from scipy.linalg import expm
-from scipy.sparse import diags, csc_matrix, csr_matrix, lil_matrix, lil_array
+from scipy.sparse import diags, csc_matrix, csr_matrix, lil_matrix, lil_array, identity
 import scipy.sparse
 from scipy.sparse.linalg import expm, expm_multiply
 
-def Lanzcos(A, _V_hat, m):
-    n = np.shape(A)[-1]
+def Lanzcos(Amult, _V_hat, m):
+    n = _V_hat.shape[0] # np.shape(A)[-1]
     alpha = np.zeros((m+1))
     beta = np.zeros((m+1))
     v = np.zeros((n, m+1))
@@ -18,9 +18,10 @@ def Lanzcos(A, _V_hat, m):
     for i in range(1, m+1):
         beta[i] = norm(v_hat[:,i])
         v[:, i] = v_hat[:, i]/beta[i]
-        alpha[i] = v[:, i].T @ A @ v[:, i]
-        v_hat[:, i+1] = A@v[:,i] - alpha[i]*v[:,i] - beta[i]*v[:,i-1]
-        
+        Av = Amult(v[:, i])
+        alpha[i] = Av.dot(v[:,i]) # v[:, i].T @ A @ v[:, i]
+        v_hat[:, i+1] = Av - alpha[i]*v[:,i] - beta[i]*v[:,i-1]
+        print(i,alpha[i],beta[i])
 
     beta = beta[2:m+1]
     alpha = alpha[1:m+1]
@@ -28,10 +29,17 @@ def Lanzcos(A, _V_hat, m):
     return diags([beta, alpha, beta], [-1,0,1]), v[:, 1:m+1], norm(_V_hat)
 
 def LanzcosExp(A, v_1, m):
-    H, V, beta = Lanzcos(A, v_1, m)
+    nv_1 = np.linalg.norm(v_1)
+    H, V, beta = Lanzcos(lambda v: A@v, v_1, m)
+    # log_nv_1 = np.log(nv_1)
+    # H, V, beta = Lanzcos(lambda v: A@v + log_nv_1*v , v_1/nv_1, m)
+    ## H, V, beta = Lanzcos(lambda v: A@v , v_1/nv_1, m)
+    ## H += log_nv_1*identity(H.shape[0]) 
     e_1 = np.zeros((m))
     e_1[0] = 1
[…]

### 2024-09-11 allenCahn.py

@@ -1,0 +1,43 @@
+import numpy as np
+from ufl import *
+from dune.ufl import Space, Constant
+from dune.fem.function import gridFunction
+
+dimR = 1
+time = Constant(0,"time")
+sourceTime = Constant(0,"sourceTime")
+domain = [0, 0], [1, 1], [60, 60]
+
+space = Space(2,dimRange=dimR)
+x,u,v,n = ( SpatialCoordinate(space), TrialFunction(space), TestFunction(space), FacetNormal(space) )
+
+eps = Constant(0.01,"eps")
+theta = Constant(0.8,"theta")
+theta_c = Constant(1.6,"theta_c")
+
+"""
+a = ( eps**2 * inner(grad(u),grad(v))
+      - theta/2 * ln( (1-u[0])/(1+u[0]) ) * v[0]
+      - theta_c * dot(u,v)
+    ) * dx
+"""
+
+a = ( eps**2 * inner(grad(u),grad(v))
+      + (dot(u,u)-1) * dot(u,v)
+    ) * dx
+
+"""
+def test1(gridView):
+    h = 0.01
+    g0 = lambda x,x0,T: conditional(x-x0<-T/2,0,conditional(x-x0>T/2,0,sin(2*pi/T*(x-x0))**3))
+    G  = lambda x,y,x0,y0,T: g0(x,x0,T)*g0(y,y0,T)
+    return a, 20, 0.9*G(x[0],x[1],0.5,0.5,50*h), None
+"""
+
+def test2(gridView):
+    tauFE = 0.8   # AC on level=0
+    @gridFunction(gridView,name="random",order=1)
+    def u0(x):
+        return 1.8*np.random.rand(1)[0]-0.9
+    np.random.seed(100)
+    return a, 24, tauFE, u0, None

### 2024-09-11 kiops.py

@@ -1,0 +1,345 @@
+# taken from https://github.com/uofs-simlab/pythOS/tree/main
+
+import math
+
+import numpy
+import scipy.linalg
+import numpy as np
+from scipy.sparse import diags, csc_matrix, csr_matrix, lil_matrix, lil_array
+import scipy.sparse
+from scipy.sparse.linalg import expm, expm_multiply
+from numpy.linalg import norm
+
+def kiops(τ_out, A, u, tol = 1e-10, m_init = 10, mmin = 10, mmax = 128, iop = 2, task1 = False):
+   """
+      kiops(tstops, A, u; kwargs...) -> (w, stats)
+
+   Evaluate a linear combinaton of the ``φ`` functions evaluated at ``tA`` acting on
+   vectors from ``u``, that is
+
+   ```math
+   w(i) = φ_0(t[i] A) u[0, :] + φ_1(t[i] A) u[1, :] + φ_2(t[i] A) u[2, :] + ...
+   ```
+
+   The size of the Krylov subspace is changed dynamically during the integration.
+   The Krylov subspace is computed using the incomplete orthogonalization method.
+
+   Arguments:
+   - `τ_out`    - Array of `τ_out`
+   - `A`        - the matrix argument of the ``φ`` functions
+   - `u`        - the matrix with rows representing the vectors to be multiplied by the ``φ`` functions
+
+   Optional arguments:
+   - `tol`      - the convergence tolerance required (default: 1e-7)
+   - `mmin`, `mmax` - let the Krylov size vary between mmin and mmax (default: 10, 128)
+   - `m`        - an estimate of the appropriate Krylov size (default: mmin)
+   - `iop`      - length of incomplete orthogonalization procedure (default: 2)
+   - `task1`     - if true, divide the result by 1/T**p
+
+   Returns:
+   - `w`      - the linear combination of the ``φ`` functions evaluated at ``tA`` acting on the vectors from ``u``
+   - `stats[0]` - number of substeps
+   - `stats[1]` - number of rejected steps
+   - `stats[2]` - number of Krylov steps
+   - `stats[3]` - number of matrix exponentials
+   - `stats[4]` - Error estimate
+   - `stats[5]` - the Krylov size of the last substep
+
+   `n` is the size of the original problem
[…]

### 2024-09-11 model_bgfix.py

@@ -1,0 +1,158 @@
+from inspect import signature
+
+import numpy
+
+from dune.ufl import cell
+from ufl import *
+from dune.common import FieldVector, comm
+from dune.grid import reader, cartesianDomain
+from dune.fem.function import uflFunction
+
+
+def parprint(*args, **kwargs):
+    if comm.rank == 0:
+        print(*args, **kwargs, flush=True)
+
+
+def model(Model, dim=2, **kwargs):
+    x = SpatialCoordinate(cell(dim))
+
+    class BGModel(Model):
+
+        # overloaded source applying bgFix
+
+        def consBG(U):
+            BG, _, _, _, _ = Model.bg(0,x)
+            return U + BG
+        def toPrimBG(U):
+            return as_vector([Model.toPrim(consBG(U))])
+        def thetaTilde(U):
+            BG, _, _, thetaBG, _ = Model.bg(0,x)
+            _,  _, _, _, theta   = Model.toPrim(U + BG)
+            return theta - thetaBG
+        def spongeLayer(x):
+            return 0
+
+        def S_e(t,x,U,DU):
+            BG, _, _, _, _ = Model.bg(t,x)
+
+            U_tot = U + BG
+            S_tot = Model.S_e(t,x, U_tot, DU)
+
+            S_bg = Model.S_e(t,x, BG, DU)
+
+            sponge = BGModel.spongeLayer(x)
+            # return (S_tot - S_bg)*(1-sponge)
+            return S_tot - S_bg - sponge*U
+
+        if hasattr(Model, "F_c"):
+            # overloaded flux applying bgFix
+            def F_c(t,x,U):
+                BG, _, _, _, _ = Model.bg(t,x)
+
+                U_tot = U + BG
+                res_tot = Model.F_c(t,x, U_tot)
+
+                res_bg = Model.F_c(t,x, BG)
+
+                sponge = BGModel.spongeLayer(x)
+                # return (res_tot - res_bg)*(1-sponge) # + res_bg*sponge
+                return res_tot - res_bg
+            # interface method needed for LLF and time step control
+            def maxWaveSpeed(t,x,U,n):
+                BG, _, _, _, _ = Model.bg(t,x)
+                U_tot = U + BG
+                return Model.maxWaveSpeed(t,x,U_tot, n)
+
+        if hasattr(Model, "F_v"):
+            # overloaded viscous flux applying bgFix
+            def F_v(t,x,U, DU):
[…]

### 2024-09-11 parabolicTest.py

@@ -1,0 +1,42 @@
+from ufl import *
+from dune.ufl import Space, Constant
+
+dimR = 1
+time = Constant(0,"time")
+sourceTime = Constant(0,"sourceTime")
+domain = [0, 0], [1, 1], [80, 80]
+
+space = Space(2,dimRange=dimR)
+x,u,v,n = ( SpatialCoordinate(space), TrialFunction(space), TestFunction(space), FacetNormal(space) )
+
+def model(exact,dtExact,nonLinear):
+    a  = ( inner(grad(u),grad(v)) + dot(nonLinear(u),v) ) * dx
+    bf = lambda t: ( ( dtExact(t)[0] - div(grad(exact(t)[0])) ) * v[0]
+                     + dot(nonLinear(exact(t)),v) 
+                   ) * dx 
+    bg = lambda t: dot(grad(exact(t)[0]),n)*v[0]*ds
+    return a-bf(sourceTime)-bg(sourceTime)
+
+def paraTest0(gridView):
+    tauFE = 7e-5 # check if this is correct
+    freq = 6*pi
+    xExact = as_vector([ cos(freq*x[0]) ])
+    exact = lambda t: exp(-freq**2*t) * xExact
+    dtExact = lambda t: -freq**2*exp(-freq**2*t) * xExact
+    return model(exact,dtExact, lambda u: as_vector([0])), 0.5, tauFE, exact(time), exact
+
+def paraTest1(gridView):
+    tauFE = 7e-5 # check if this is correct
+    freq = 6*pi
+    xExact = as_vector([ cos(freq*x[0]) ])
+    exact = lambda t: exp(-freq**2*t) * xExact
+    dtExact = lambda t: -freq**2*exp(-freq**2*t) * xExact
+    return model(exact,dtExact, lambda u: (1+dot(u,u))*u), 0.5, tauFE, exact(0), exact
+
+def paraTest2(gridView):
+    tauFE = 7e-5 # check if this is correct
+    freq = 6*pi
+    xExact = as_vector([ cos(freq*x[0]) ])
+    exact = lambda t: cos(freq**2*t) * xExact
+    dtExact = lambda t: -freq**2*sin(freq**2*t) * xExact
+    return model(exact,dtExact, lambda u: (1+dot(u,u))*u), 0.5, tauFE, exact(time), exact

### 2024-09-11 plot_data.py

@@ -7,9 +7,8 @@ import pickle as pickle
 e_threshold = 1e-10
 
 # %%
-with open("Experiment_Data.pickle", "rb") as file:
-    data = pickle.load(file)
-ns = [1000, 10000, 100000,1000000]
+data = pd.read_csv("Experiment_Data.csv")
+ns = data["N"].unique()
 # %%
 for n in ns:
     ndata = data[data["N"] == n]
@@ -22,7 +21,7 @@ for n in ns:
             mean_Error = np.mean(methoddata["Error"])
         else:
             plt.loglog(methoddata["M"], methoddata["Error"], label= method)
-        plt.xlabel("M")
+        plt.xlabel("Dimention of Kyrlov subspace $m$")
         plt.ylabel("Error")
         plt.title("Graph Compairing M and Error for N={0}".format(n))
         plt.legend()
@@ -30,17 +29,17 @@ for n in ns:
         plt.figure(2)
         if method == "Scipy":
             mean_computation_time = np.mean(methoddata["Computation Time"])
-            plt.hlines(mean_computation_time, np.min(ndata["M"]), np.max(ndata["M"]), colors="r", linestyles="dashed")
+            plt.hlines(mean_computation_time, np.min(ndata["M"]), np.max(ndata["M"]), colors="r", linestyles="dashed", label="Scipy")
         else:
             plt.loglog(methoddata["M"], methoddata["Computation Time"], label= method)
-        plt.xlabel("M")
+        plt.xlabel("Dimention of Kyrlov subspace $m$")
         plt.ylabel("Computation Time $s$")
         plt.title("Graph Compairing M and Computation Time for N={0}".format(n))
         plt.legend()
         
         plt.figure(3)
         if method == "Scipy":
-            plt.vlines(mean_computation_time, np.min(ndata["Error"]), np.max(ndata["Error"]), colors="r", linestyles="dashed")
+            plt.vlines(mean_computation_time, np.min(ndata["Error"]), np.max(ndata["Error"]), colors="r", linestyles="dashed", label="Scipy")
         else:
             plt.loglog(methoddata["Computation Time"], methoddata["Error"], label= method)
         plt.xlabel("Computation Time $s$")
@@ -61,13 +60,12 @@ for n in ns:
 # %% Plotting time for error to be bellow given bound
 cut_data = data[data["Error"] < e_threshold]
 ns = cut_data["N"].unique()
-for method in cut_data["Method"].unique():
+for method in data["Method"].unique():
[…]

### 2024-09-11 produce_data.py

@@ -9,6 +9,7 @@ import scipy.sparse
 from Arnoldi import ArnoldiExp
 from Stable_Lanzcos import LanzcosExp
 from NBLA import NBLAExp
+from kiops import KiopsExp
 
 # Collects data on performance of different methods
 # Independant Variables:
@@ -21,22 +22,19 @@ from NBLA import NBLAExp
 def GetA(n):
     A = diags([-1,2,-1], [-1,0,1], shape=(n,n))
     A = csc_matrix(A)
-    return A
+    return A / n**2
 
 methods = {
     "Scipy": lambda A, v, m: expm_multiply(A,v),
     "Arnoldi": ArnoldiExp,
     "Lanzcos": LanzcosExp#,
+    #"Kiops": lambda A, V, m: KiopsExp(A,V)
     #"NBLA": NBLAExp
 }
 m = [1,2,3,4,5,6,7,8,9,10,20,30,40,50,60,70,80,90,100]
-n = [10,20,30,40,50,60,70,80,90,
-     100,200,300,400,500,600,700,800,900,
-     1000,2000,3000,4000,5000,6000,7000,8000,9000,
-     10000,20000,30000,40000,50000,60000,70000,80000,90000,
-     100000,200000,300000,400000,500000,600000,700000,800000,900000,
-     1000000
-     ]
+n = [2**i for i in range(0, 8)]
+print(n)
+
 v = np.random.rand(np.max(n))
 method_names = list(methods.keys())
 mlen = len(m)
@@ -61,6 +59,7 @@ computation_times = []
 true_values = []
 scipy_computing_time =[]
 for N in n:
+    print("Computing {0}".format(N))
     A = GetA(N)
     V = v[0:N]
     start = time.time()
@@ -108,6 +107,5 @@ data["Computation Time"] = computation_times
 
 data.dropna()
 
-with open("Experiment_Data.pickle", "wb") as file:
-    pickle.dump(data, file, pickle.HIGHEST_PROTOCOL)
+data.to_csv("Experiment_Data.csv")
 

### 2024-09-11 risingbubble.py

@@ -1,0 +1,138 @@
+import sys
+
+from dune.fem.function import gridFunction
+from dune.ufl import cell, Constant
+from ufl import SpatialCoordinate, sqrt, exp, sin, pi, conditional, as_vector
+
+from model_euler_atmos import model as AtmosphericModel
+from model_bgfix import model as BgFixModel
+
+
+def RisingBubble(dim=2):
+    Model = AtmosphericModel(dim=dim)
+    Model.atmos = "neutral"
+    Model.hSpeed0 = 0.
+    Model.T0 = 303.15
+    #Model.K = 0.1
+
+    x = SpatialCoordinate(cell(dim))
+    Model.dx = Constant(0., name="dx")
+
+    z   = x[dim-1] # z-coordinate
+    #x_c = 533.     # x-center of perturbation ball
+    x_c = 500. + 0.5*Model.dx  # x-center of perturbation ball
+    z_c = 520.     # z-center of perturbation ball
+    a   = 50       # radius of perturbation ball
+
+    r = sqrt((x[0] - x_c)**2 + (z - z_c)**2)
+
+    BG, p, T, theta, v = Model.bg(0, x)
+
+    A = 0.5
+    s = 100.
+
+    deltaTheta = A * conditional(r <= a,
+        1.0,
+        exp( -(r-a)**2/s**2)
+    )
+    newTheta = theta + deltaTheta
+
+    # introduce pot. temp. to conservative variables keeping
+    # the pressure equal to the background pressure
+
+    rho = pow( Model.p0, Model.kappa ) * pow( p, 1./Model.gamma ) / Model.R_d / newTheta
+
+    Model.U0 = Model.toCons(as_vector([rho] + v + [newTheta]))
+
+    Model.boundary = {1: lambda t,x,U: Model.reflect(U,[-1, 0]),
+                      2: lambda t,x,U: Model.reflect(U,[ 1, 0]),
+                      3: lambda t,x,U: Model.reflect(U,[ 0,-1]),
+                      4: lambda t,x,U: Model.reflect(U,[ 0, 1])}
+
+
+    Model.domain = [0]*dim, [1000]*(dim-1)+[2000], [40]*(dim-1)+[80]
+    Model.endTime = 400 # 1200.
+    Model.name = "RisingBubble"
+
+    return BgFixModel(Model, dim)
+
+import sys
+import numpy as np
+import matplotlib.pyplot as plt
+from dune.grid import structuredGrid
[…]

### 2024-09-11 steppers.py

@@ -5,29 +5,45 @@
 # ## Setup
 
 # %%
+import sys
+
 import matplotlib.pyplot as plt
 import numpy as np
-from scipy.sparse.linalg import cg, LinearOperator
+from scipy.linalg import norm
+from scipy.sparse.linalg import lgmres, gmres, LinearOperator, aslinearoperator
 from scipy.sparse import identity, diags
-
-from dune.grid import structuredGrid as leafGridView
-from dune.fem.space import lagrange
-from dune.fem import integrate
-from dune.fem.operator import galerkin
-from ufl import ( TestFunction, TrialFunction, SpatialCoordinate, FacetNormal,
-                  dx, ds, div, grad, dot, inner, conditional, as_vector,
-                  exp, sin, pi )
+from scipy.optimize import newton_krylov, KrylovJacobian
+import random
 
 from scipy.sparse.linalg import expm_multiply
+expm_sci = [lambda A,x,m: expm_multiply(A,x),"Scipy"]
+from Stable_Lanzcos import LanzcosExp
+expm_lanzcos = [lambda A,x,m: LanzcosExp(A,x,m),"Lanzcos"]
+from NBLA import NBLAExp
+expm_nbla = [lambda A,x,m: NBLAExp(A,x,m),"NBLA"]
+from Arnoldi import ArnoldiExp 
+expm_arnoldi = [lambda A,x,m: ArnoldiExp(A,x,m),"Arnoldi"]
+from kiops import KiopsExp
+expm_kiops = [lambda A,x,m: KiopsExp(A,x),"Kiops"]
+
+from dune.grid import cartesianDomain
+from dune.alugrid import aluCubeGrid as leafGridView
+from dune.fem.space import lagrange
+from dune.fem import integrate, threading, globalRefine
+from dune.fem.view import adaptiveLeafGridView as view
+from dune.fem.operator import galerkin
+from dune.fem.function import gridFunction
+from ufl import dx, dot, inner, grad, TestFunction, TrialFunction
+
 # %% [markdown]
 #
 # Simple utility function to show the result of a simulation
 
 # %%
-def printResult(method,error):
-    print(method,"L^2, H^1 error:",'{:0.5e}, {:0.5e}'.format(
+def printResult(time,error,*args):
+    print(time,'{:0.5e}, {:0.5e}'.format(
                  *[ np.sqrt(e) for e in integrate([error**2,inner(grad(error),grad(error))]) ]),
[…]

### 2024-09-11 article_4.tex

--- 2024-08-19 article_4.tex
@@
 \usepackage{amsmath}
+\usepackage{amssymb}
+\usepackage{algorithm}
+\usepackage[noend]{algpseudocode}
+\usepackage{float}
+\usepackage{graphicx}
+\usepackage{afterpage}
 \begin{document}
@@
 %Statement of ODE
-It is common within applied mathematics to encounter ODEs of the form
+In many simulations, it is common to encounter ODEs of the form
 \begin{align}
-\dot u(t) = Au(t) + N(u) \label{ODE}
+\dot u(t) = Au(t) + R(u) \label{ODE}
 \end{align}
-Where $A$ is a large sparse matrix and $N$ denotes some nonlinear term.
+Where $A$ is a large sparse matrix and $R$ denotes some nonlinear term.
 As a result, finding accurate approximate solutions to these equations is of major importance.\\
@@
 %Method of lines
-These ODEs also occure when solving PDEs numerically using the method of lines.
+Often ODEs also occure when solving PDEs numerically using the method of lines.
 This arrives from discretising the PDE in space yielding an ODE with respect to time, with the form shown in \eqref{ODE}.\\
@@
 \begin{align*}
-u(t+h) &= u(t) + h\dot u(t) = u(t) + h(Au(t) + N(u(t)))
+u(t+\tau) &\approx u(t) + \tau\dot u(t) = u(t) + \tau(Au(t) + N(u(t)))
 \end{align*}
@@
 \begin{align*}
-u(t+h) &= e^{Ah}u(t) + \int_0^h e^{A(h-z)}N(u(t+z)) dz \label{ODE Solution}\\
-\text{where } e^{A} &= \sum^{\inf}_{i = 0}\frac{A^n}{n!}
+u(t+\tau) &= e^{Ah}u(t) + \int_0^h e^{A(\tau-z)}N(u(t+z)) dz \label{ODE Solution}\\
+&\text{using a simple explicit approach leads to}\\
+&\approx e^{Ah}(u(t) + N(u(t))\\
+\text{where } e^{A} &= \sum^{\infty}_{i = 0}\frac{A^n}{n!}
 \end{align*}
@@
 Another class of methods known as Krylov subspace methods have gained traction recently\cite{Moler2003}.
-These work by projecting the problem onto a space with fewer dimensions (the kyrlov subspace), computing the matrix exponential and then transforming back to the original space.
-As the matrix exponential will be computed in the lower dimension kyrlov subspace the computation cost of this step will not be significant.
+These work by projecting the problem onto a space with fewer dimensions (the Kyrlov subspace), computing the matrix exponential and then transforming back to the original space.
+As the matrix exponential will be computed in the lower dimension Kyrlov subspace the computation cost of this step will not be significant.
+\section{Krylov Subspace Methods}
+In this section, we will look into existing Krylov subspace methods and then compare how these methods perform for approximating $e^{A}v$.
 
+\subsection{Alogrithms}
+The algorithms take a matrix $A\in \mathbb{R}^{n\times n}$, a vector $v \in \mathbb{R}^n$ and an integer $m$ that determines the number of dimensions of the Krylov subspace used.
+From these algorithms we will get a matrix $H \in \mathbb{R}^{m\times m}$ and another matrix $V \in \mathbb{R}^{n\times m}$ such that $A \approx VHV^T$ and $VV^T = I$.
+From here we get $Av \approx VHV^Tv = VH||v||e_1$.\\
+We can apply this to $e^A$ as follows:
+\begin{align*}
+e^A &= \sum^{\infty}_{i=0}\frac{A^i}{i\!}\\
+&= \sum^{\infty}_{i=0}\frac{(VHV^T)^i}{i\!} \\
+&= \sum^{\infty}_{i=0}\frac{VH^iV^T}{i\!} \\
+&\text {and then when computing $e^Av$ we get}\\
+e^Av &= (\sum^{\infty}_{i=0}\frac{VH^iV^T}{i\!})v \\
+&= \sum^{\infty}_{i=0}\frac{VH^iV^Tv}{i\!} \\
+&= \sum^{\infty}_{i=0}\frac{VH^i||v||e_1}{i\!} \\
+&= V(\sum^{\infty}_{i=0}\frac{H^i}{i\!})||v||e_1 \\
+&= Ve^H||v||e_1
+\end{align*}
+
+Here we give the algorithms for the 2 Krylov subspace methods we will be comparing.
+We begin with the Arnoldi algorithm.
+
+\begin{algorithm}[H]
+\caption{Arnoldi \cite{Fan2018}} %find better citation
+\begin{algorithmic}
+\Procedure{Arnoldi}{$A, \hat v_1,m$}
+\State $v_0 \gets 0$
+\For{$j = 1,2,...,m$}	
+\For{$i = 1,2,...,j$}
+\State$h_{ij} \gets v_i^T A v_i$
+\EndFor
+\State$\theta_j \gets Av_j - \sum^j_{i=1} h_{ij}v_i$
+\State$h_{j+1,j} \gets ||\theta_j||$
+\State$v_{j+1} \gets \theta_j/h_{j+1,j}$
+\EndFor
+\EndProcedure
+\end{algorithmic}
+\end{algorithm}
+Here V is given by $v_1,...,v_m$ and H is give by $h_1,...,h_m$.\\
+The algorithm below is the Lanczos algorithm and it requires a symmetric matrix. \cite{Moler2003}
+\begin{algorithm}[H]
+\caption{Lanczos \cite{OJALVO1970}}
+\begin{algorithmic}
+\Procedure{Lanczos}{$A$ symetric$, \hat v_1,m$}
+\State $v_0 = 0$
+\For{$i = 1,2,...,m$}	
+\State$\beta_i \gets || \hat v_i ||$
+\State$v_i \gets \hat v_i / || \hat v_i ||$
+\State$\alpha_i \gets v_i^T A v_i$
+\State$\hat v_{i+1} \gets Av_i - \alpha_iv_i - \beta_iv_{i-1}$
+\EndFor
+\EndProcedure
+\end{algorithmic}
+\end{algorithm}
+Here $V$ is given by ${v_1,...,v_m}$ and $H$ is tridiagonal with the leading diagonal being $\alpha_1, ..., \alpha_m$ and the upper and lower diagonals being $\beta_2,...,\beta_m$.
+
+\subsection{Numerical Results}
+Here we run numerical experiments, computing $e^Av$ for a tridiagonal matrix $A$ of sizes $131072$ and $524288$, where $A$ is a triagonal matrix with $2n^2$ along the diagonal and $-n^2$ along the upper and lower diagonals.
+We will compare the methods with respect to computation time, error and size of the Kyrlov subspace.
+Here, the error is computed as the Euclidian distance between the Scipy method and the Krlob subspace method that is being tested.
+
+\begin{figure}[H]
+    \centering
+    \begin{minipage}{0.5\textwidth}
+       \centering
+	  \includegraphics[width=\linewidth]{Plots/M v E Results for N=131072.png}
+	  \label{fig:MEe7}
+       \centering
+	  \includegraphics[width=\linewidth]{Plots/M v Comp Time Results for N=131072.png}
+	  \label{fig:MEe7}
+       \centering
+	  \includegraphics[width=\linewidth]{Plots/Comp Time v E Results for N=131072.png}
+	  \label{fig:MEe7}
+    \end{minipage}\hfill
+    \begin{minipage}{0.5\textwidth}
+       \centering
+	  \includegraphics[width=\linewidth]{Plots/M v E Results for N=524288.png}
+	  \label{fig:MEe6}
+       \centering
+	  \includegraphics[width=\linewidth]{Plots/M v Comp Time Results for N=524288.png}
+	  \label{fig:MEe6}
+       \centering
+	  \includegraphics[width=\linewidth]{Plots/Compt time v E Results for N=524288.png}
+	  \label{fig:MEe6}
+    \end{minipage}
+\end{figure}
+These results show that the Lanzcos algorithm outperforms the Arnoldi algorithm in both convergence rate and computational time.
+
+Below we compare the time required for the different algorithms to get below an error of $10^{10}$ when computing $e^Av$ for various matrix sizes.
+\begin{figure}[H]
+\includegraphics[width=\linewidth]{Plots/time to get below an error of 1e-10.png}
+\end{figure}
+Again we continue to see that the Lanzcos algorithm outperforms the Arnodli algorithm.
+\clearpage
 \newpage

### 2024-09-11 plot_data.py
Updated Plotting of data

@@ -4,7 +4,7 @@ import pandas as pd
 import matplotlib.pyplot as plt
 import pickle as pickle
 
-e_threshold = 1e-10
+e_thresholds = [1e-6, 1e-10]
 
 # %%
 data = pd.read_csv("Experiment_Data.csv")
@@ -49,37 +49,42 @@ for n in ns:
     
     plt.figure(1)
     plt.savefig("Plots/M v E Results for N={0}".format(n))
+    plt.grid(True, which="both")
     plt.close()
     plt.figure(2)
-    plt.savefig("Plots/M v Comp time Results for N={0}".format(n)) 
+    plt.grid(True, which="both")
+    plt.savefig("Plots/M v Comp Time Results for N={0}".format(n)) 
     plt.close()
     plt.figure(3)
-    plt.savefig("Plots/Compt time v E Results for N={0}".format(n))
+    plt.grid(True, which="both")
+    plt.savefig("Plots/Compt Time v E Results for N={0}".format(n))
     plt.close()
 
 # %% Plotting time for error to be bellow given bound
-cut_data = data[data["Error"] < e_threshold]
-ns = cut_data["N"].unique()
-for method in data["Method"].unique():
-    method_data = cut_data[cut_data["Method"] == method]
-    result = []
-    for n in ns:
-        ndata = method_data[method_data["N"] == n]
-        if method == "Scipy":
-            continue
+for e_threshold in e_thresholds:
+    cut_data = data[data["Error"] < e_threshold]
+    ns = cut_data["N"].unique()
+    for method in data["Method"].unique():
+        method_data = cut_data[cut_data["Method"] == method]
+        result = []
+        for n in ns:
+            ndata = method_data[method_data["N"] == n]
+            if method == "Scipy":
+                continue
+            try:
+                result.append(ndata["Computation Time"].to_list()[ndata["Error"].argmin()])
+            except:
+                result.append(None)
         try:
-            result.append(ndata["Computation Time"].to_list()[ndata["Error"].argmin()])
+            plt.loglog(ns, result, label = method)
         except:
-            result.append(None)
-    try:
-        plt.loglog(ns, result, label = method)
-    except:
-        continue
-plt.legend()
-plt.title("A graph Showing the time to get below an error of {0} for different matrix sizes".format(e_threshold))
-plt.xlabel("N")
-plt.ylabel("Computation Time")
-plt.savefig("Plots/time to get below an error of {0}.png".format(e_threshold))
[…]

### 2024-09-11 produce_data.py

@@ -32,7 +32,7 @@ methods = {
     #"NBLA": NBLAExp
 }
 m = [1,2,3,4,5,6,7,8,9,10,20,30,40,50,60,70,80,90,100]
-n = [2**i for i in range(0, 8)]
+n = [2**i for i in range(0, 20)]
 print(n)
 
 v = np.random.rand(np.max(n))

### 2024-09-11 plot_data.py
Graphs Neetened

@@ -48,8 +48,8 @@ for n in ns:
         plt.legend()
     
     plt.figure(1)
-    plt.savefig("Plots/M v E Results for N={0}".format(n))
     plt.grid(True, which="both")
+    plt.savefig("Plots/M v E Results for N={0}".format(n))
     plt.close()
     plt.figure(2)
     plt.grid(True, which="both")
@@ -85,6 +85,7 @@ for e_threshold in e_thresholds:
     plt.ylabel("Computation Time")
     plt.grid(True, which="both")
     plt.savefig("Plots/time to get below an error of {0}.png".format(e_threshold))
+    plt.close()
 
 
 

### 2024-09-11 article_4.tex
Added Graphs to Report

@@
        \centering
-	  \includegraphics[width=\linewidth]{Plots/Compt time v E Results for N=524288.png}
+	  \includegraphics[width=\linewidth]{Plots/Comp time v E Results for N=524288.png}
 	  \label{fig:MEe6}
@@
 
-Below we compare the time required for the different algorithms to get below an error of $10^{10}$ when computing $e^Av$ for various matrix sizes.
+Below we compare the time required for the different algorithms to get below an error of $10^6$ and $10^{10}$ when computing $e^Av$ for various matrix sizes.
+\begin{figure}[H]
+\includegraphics[width=\linewidth]{Plots/time to get below an error of 1e-06.png}
+\end{figure}
 \begin{figure}[H]

### 2024-09-11 plot_data.py
Fixed Graph Fomatting

@@ -57,7 +57,7 @@ for n in ns:
     plt.close()
     plt.figure(3)
     plt.grid(True, which="both")
-    plt.savefig("Plots/Compt Time v E Results for N={0}".format(n))
+    plt.savefig("Plots/Comp Time v E Results for N={0}".format(n))
     plt.close()
 
 # %% Plotting time for error to be bellow given bound
@@ -80,7 +80,7 @@ for e_threshold in e_thresholds:
         except:
             continue
     plt.legend()
-    plt.title("A graph Showing the time to get below an error of {0} for different matrix sizes".format(e_threshold))
+    plt.title("Graph Showing time to get below error of {0} for different matrix sizes".format(e_threshold))
     plt.xlabel("N")
     plt.ylabel("Computation Time")
     plt.grid(True, which="both")

### 2024-09-12 article_4.tex
Spelling Corrections

@@
 \subsection{Numerical Results}
-Here we run numerical experiments, computing $e^Av$ for a tridiagonal matrix $A$ of sizes $131072$ and $524288$, where $A$ is a triagonal matrix with $2n^2$ along the diagonal and $-n^2$ along the upper and lower diagonals.
+Here we run numerical experiments, computing $e^Av$ for a tridiagonal matrix $A$ of sizes $131072$ and $524288$, where $A$ is a tridiagonal matrix with $2n^2$ along the diagonal and $-n^2$ along the upper and lower diagonals.
 We will compare the methods with respect to computation time, error and size of the Kyrlov subspace.
-Here, the error is computed as the Euclidian distance between the Scipy method and the Krlob subspace method that is being tested.
+The error is computed as the Euclidian distance between Scipy and the Krlov subspace method being tested.
 

### 2024-09-19 FETester.py
Fixed Hash Bug and error bug

@@ -2,6 +2,7 @@ import numpy as np
 import pickle as pickle
 import os as os
 import hashlib
+import copy
 from ufl import *
 from dune.ufl import Space, Constant
 from dune.fem import integrate
@@ -76,8 +77,10 @@ class Tester():
 
         # Generate test stepper data if it doesn't exist
         test_stepper = test_stepper(op, **stepper_args)
+        temp = copy.deepcopy(stepper_args)
+        temp['exp_v'][0] = 0
         test_stepper_name = test_stepper.name +\
-                            hashlib.sha1( repr(sorted(stepper_args.items())).encode('utf-8') ).hexdigest()
+                            hashlib.sha1( repr(sorted(temp.items())).encode('utf-8') ).hexdigest()
         test_file_name = self.get_test_results_file(tau, test_stepper_name, end_time)
         if not os.path.isfile(test_file_name):
             self.test_results, self.test_countN = self.run(tau, test_stepper, self.initial_condition, self.seed_time, end_time)
@@ -143,7 +146,7 @@ if __name__ == "__main__":
             #     exact_error = ...
 
             # write file self.test_results.plot()
-            results += [ ["EXPARN",gridView.size(0),tau,error.value,
+            results += [ ["EXPARN",gridView.size(0),tau,H1err[0],
                                  tester.target_countN,tester.test_countN] ]
 
     # produce plots using 'results'

### 2024-09-19 FETester.py
Fixed bug with seed not working properly

@@ -3,6 +3,8 @@ import pickle as pickle
 import os as os
 import hashlib
 import copy
+import pandas as pd
+import matplotlib.pyplot as plt
 from ufl import *
 from dune.ufl import Space, Constant
 from dune.fem import integrate
@@ -53,10 +55,10 @@ class Tester():
 
         # If not generate it
         else:
-            time = Constant(0,"time")
-            initial_condition, _ = self.run(tau, setup_stepper, initial_condition, 0, self.seed_time)
+            self.initial_condition, _ = self.run(tau, setup_stepper, initial_condition, 0, self.seed_time)
             with open(self.intitial_file_name, 'wb') as file:
-                pickle.dump(initial_condition.as_numpy[:], file)
+                pickle.dump(self.initial_condition.as_numpy[:], file)
+        self.initial_condition.plot()
     
     def run_test(self, tau, test_stepper, stepper_args, end_time):
         
@@ -95,7 +97,7 @@ class Tester():
     def run(self, tau, stepper, initial_condition, start_time, end_time):
         # Runs for a given stepper
         current_step = initial_condition.copy()
-        time = Constant(self.seed_time)
+        time = Constant(start_time)
         countN = 0
         while time.value < end_time:
             stepper(target=current_step, tau = tau)
@@ -105,6 +107,7 @@ class Tester():
 
     def produce_results(self, tau, stepper, stepper_args, end_time):
         self.run_test(tau, stepper, stepper_args, end_time)
+        self.target.plot()
 
 if __name__ == "__main__":
     from allenCahn import dimR, time, sourceTime, domain
@@ -112,8 +115,12 @@ if __name__ == "__main__":
 
     results = []
 
-    for N in [15,30,60]:
-        for tau in [1e-2,4e-2,1.6e-1]:
+    start_time = 3
+    end_time = 3
+
+    for N in [2**i for i in range(0, 1)]:
+        N = 25
+        for tau in [1e-2]:
             temp = list(domain)
             temp[2] = [N,N]
             domain = temp
@@ -134,9 +141,8 @@ if __name__ == "__main__":
                 m = 5
                 args["expv_args"] = {"m":m}
 
-            end_time = 3
 
-            tester = Tester(u_h, op, "Allen Cahn", 1)
+            tester = Tester(u_h, op, "Allen Cahn", start_time)
             
             tester.produce_results(tau, exp_arnoldi_stepper, args, end_time)
 
[…]

### 2024-09-19 FETester.py
Removed Plotting

@@ -58,7 +58,6 @@ class Tester():
             self.initial_condition, _ = self.run(tau, setup_stepper, initial_condition, 0, self.seed_time)
             with open(self.intitial_file_name, 'wb') as file:
                 pickle.dump(self.initial_condition.as_numpy[:], file)
-        self.initial_condition.plot()
     
     def run_test(self, tau, test_stepper, stepper_args, end_time):
         
@@ -107,7 +106,6 @@ class Tester():
 
     def produce_results(self, tau, stepper, stepper_args, end_time):
         self.run_test(tau, stepper, stepper_args, end_time)
-        self.target.plot()
 
 if __name__ == "__main__":
     from allenCahn import dimR, time, sourceTime, domain
@@ -115,11 +113,10 @@ if __name__ == "__main__":
 
     results = []
 
-    start_time = 3
+    start_time = 1
     end_time = 3
 
-    for N in [2**i for i in range(0, 1)]:
-        N = 25
+    for N in [2**i for i in range(4, 8)]:
         for tau in [1e-2]:
             temp = list(domain)
             temp[2] = [N,N]
@@ -130,8 +127,6 @@ if __name__ == "__main__":
 
             model, T, tauFE, u0, exact = problem(gridView)
             op = galerkin(model, domainSpace=space, rangeSpace=space)
-
-            tau = 1e-2
 
             u_h = space.interpolate(u0, name='u_h')
 

### 2024-09-20 FETester.py
Plotting Results

@@ -116,44 +116,62 @@ if __name__ == "__main__":
     start_time = 1
     end_time = 3
 
-    for N in [2**i for i in range(4, 8)]:
-        for tau in [1e-2]:
-            temp = list(domain)
-            temp[2] = [N,N]
-            domain = temp
+    exp_methods = ["EXPARN", "EXPLAN", "EXPKIOPS"]
 
-            gridView = view(leafGridView(cartesianDomain(*domain)) )
-            space = lagrange(gridView, order=1, dimRange=dimR)
+    for exp_method in exp_methods:
+        print("EXP method:{0}".format(exp_method))
+        for N in [2**i for i in range(3, 8)]:
+            print("N:{0}".format(N))
+            for tau in [1/2**i for i in range(3, 8)]:
+                print("Tau:{0}".format(tau))
 
-            model, T, tauFE, u0, exact = problem(gridView)
-            op = galerkin(model, domainSpace=space, rangeSpace=space)
+                temp = list(domain)
+                temp[2] = [N,N]
+                domain = temp
 
-            u_h = space.interpolate(u0, name='u_h')
+                gridView = view(leafGridView(cartesianDomain(*domain)) )
+                space = lagrange(gridView, order=1, dimRange=dimR)
 
-            exp_arnoldi_stepper, args = steppersDict["EXPARN"]
+                model, T, tauFE, u0, exact = problem(gridView)
+                op = galerkin(model, domainSpace=space, rangeSpace=space)
 
-            if "exp_v" in args.keys():
-                m = 5
-                args["expv_args"] = {"m":m}
+                u_h = space.interpolate(u0, name='u_h')
+
+                exp_stepper, args = steppersDict[exp_method]
+                if "exp_v" in args.keys():
+                    m = 5
+                    args["expv_args"] = {"m":m}
 
 
-            tester = Tester(u_h, op, "Allen Cahn", start_time)
-            
-            tester.produce_results(tau, exp_arnoldi_stepper, args, end_time)
+                tester = Tester(u_h, op, "Allen Cahn", start_time)
+                
+                tester.produce_results(tau, exp_stepper, args, end_time)
 
-            error = tester.test_results - tester.target
-            H1err = [ np.sqrt(e) for e in integrate([error**2,inner(grad(error),grad(error))]) ]
-            # if exact is not None:
-            #     exact_error = ...
+                error = tester.test_results - tester.target
[…]

### 2024-09-20 Arnoldi.py
fix some issues and allow for problems with exact solution in Tester

@@ -15,17 +15,18 @@ def Arnoldi(A, v_1, m):
             h[i,j] = np.dot(w, v[:,i])
             w -= h[i,j]*v[:,i]
             
-        if j < m:
-            h[j+1,j] = norm(w)
-            v[:,j+1] = w/h[j+1,j]
-        else:
+        h[j+1,j] = norm(w)
+        if h[j+1,j] < 1e-10:
             break
+        v[:,j+1] = w/h[j+1,j]
 
-    return csc_matrix(h[0:m,0:m]), csc_matrix(v[:,0:m])
+    return csc_matrix(h[0:j,0:j]), csc_matrix(v[:,0:j])
 
 def ArnoldiExp(A, v_1, m):
     H, V = Arnoldi(A, v_1, m)
-    e_1 = np.zeros((m))
+    if H.shape[0] == 0:
+        return np.zeros_like(v_1)
+    e_1 = np.zeros((H.shape[0]))
     e_1[0] = 1
     ret = V@expm_multiply(H,e_1)*np.linalg.norm(v_1)
     # print(v_1.dot(v_1), ret.dot(ret))

### 2024-09-20 FETester.py

@@ -16,6 +16,9 @@ from dune.alugrid import aluCubeGrid as leafGridView
 from dune.fem.view import adaptiveLeafGridView as view
 from steppers import BEStepper, steppersDict
 
+# results_folder = "FETests"
+results_folder = "FETests2009"
+
 class Tester():
     def __init__(self, 
                  initial_condition, 
@@ -24,6 +27,7 @@ class Tester():
                  seed_time = 0, 
                  setup_tau = 1e-2,
                  setup_stepper = BEStepper,
+                 exact = None,
                  **stepper_args):
 
         self.op = op
@@ -31,9 +35,10 @@ class Tester():
         self.initial_condition = initial_condition
         self.seed_time = seed_time
         self.setup_stepper = setup_stepper(op, **stepper_args)
-        self.folder = "FETests/Problem:{0}_Grid_Size:{4}_Setup_Tau:{2}_Setup_Stepper:{1}_Seed_Time:{3}".format(problem_name,self.setup_stepper.name,setup_tau,self.seed_time,self.N)
+        self.folder = results_folder + "/Problem:{0}_Grid_Size:{4}_Setup_Tau:{2}_Setup_Stepper:{1}_Seed_Time:{3}".format(problem_name,self.setup_stepper.name,setup_tau,self.seed_time,self.N)
+        self.exact = exact
         
-        self.run_setup(setup_tau, self.setup_stepper, initial_condition)
+        self.run_setup(setup_tau, initial_condition)
 
     def get_initial_file(self):
         return self.folder + "_Initial.pickle"
@@ -44,7 +49,10 @@ class Tester():
     def get_test_results_file(self, tau, stepper_name, end_time):
         return self.folder + "_Tau:{0}_End_Time:{1}_Stepper:{2}.pickle".format(tau, end_time, stepper_name)
 
-    def run_setup(self, tau, setup_stepper, initial_condition):
+    def run_setup(self, tau, initial_condition):
+        if self.exact is not None:
+            return
+
         # File path to initial data
         self.intitial_file_name = self.get_initial_file()
 
@@ -55,26 +63,34 @@ class Tester():
 
         # If not generate it
         else:
-            self.initial_condition, _ = self.run(tau, setup_stepper, initial_condition, 0, self.seed_time)
+            self.initial_condition, _ = self.run(tau, self.setup_stepper, initial_condition, 0, self.seed_time)
             with open(self.intitial_file_name, 'wb') as file:
[…]

### 2024-09-20 steppers.py

@@ -231,7 +231,7 @@ class BEStepper(BaseStepper):
 
     def __call__(self, target, tau):
         try:
-            self.N.model.sourceTime += time.value
+            self.N.model.sourceTime += tau
         except AttributeError:
             pass
         self.setup(target,tau)

### 2024-09-21 article_4.tex
Spelling Corrections

@@
 %Method of lines
-Often ODEs also occure when solving PDEs numerically using the method of lines.
+Often ODEs also occur when solving PDEs numerically using the method of lines.
 This arrives from discretising the PDE in space yielding an ODE with respect to time, with the form shown in \eqref{ODE}.\\
 %ODE Solvers
-Soving these ODEs numerically works by progressively stepping through time such as with the explicit forward Euler method, which is given as:
+Solving these ODEs numerically works by progressively stepping through time such as with the explicit forward Euler method, which is given as:
 \begin{align*}
@@
        \centering
-	  \includegraphics[width=\linewidth]{Plots/Comp time v E Results for N=524288.png}
+	  \includegraphics[width=\linewidth]{Plots/Comp Time v E Results for N=524288.png}
 	  \label{fig:MEe6}

### 2024-09-23 FETester.py
Plotting Results

@@ -5,6 +5,7 @@ import hashlib
 import copy
 import pandas as pd
 import matplotlib.pyplot as plt
+import sys
 from ufl import *
 from dune.ufl import Space, Constant
 from dune.fem import integrate
@@ -16,8 +17,8 @@ from dune.alugrid import aluCubeGrid as leafGridView
 from dune.fem.view import adaptiveLeafGridView as view
 from steppers import BEStepper, steppersDict
 
-# results_folder = "FETests"
-results_folder = "FETests2009"
+results_folder = "FETests"
+# results_folder = "FETests2009"
 
 class Tester():
     def __init__(self, 
@@ -26,6 +27,7 @@ class Tester():
                  problem_name, 
                  seed_time = 0, 
                  setup_tau = 1e-2,
+                 target_tau = 1e-4,
                  setup_stepper = BEStepper,
                  exact = None,
                  **stepper_args):
@@ -35,6 +37,7 @@ class Tester():
         self.initial_condition = initial_condition
         self.seed_time = seed_time
         self.setup_stepper = setup_stepper(op, **stepper_args)
+        self.target_tau = target_tau
         self.folder = results_folder + "/Problem:{0}_Grid_Size:{4}_Setup_Tau:{2}_Setup_Stepper:{1}_Seed_Time:{3}".format(problem_name,self.setup_stepper.name,setup_tau,self.seed_time,self.N)
         self.exact = exact
         
@@ -78,9 +81,9 @@ class Tester():
 
         # Generate target data if it doesn't exist
         if self.exact is None:
-            target_file_name = self.get_target_results_file(tau, end_time)
+            target_file_name = self.get_target_results_file(self.target_tau, end_time)
             if not os.path.isfile(target_file_name):
-                self.target,self.target_countN = self.run(tau, self.setup_stepper, self.initial_condition, self.seed_time, end_time)
+                self.target,self.target_countN = self.run(self.target_tau, self.setup_stepper, self.initial_condition, self.seed_time, end_time)
                 with open(target_file_name, 'wb') as file:
                     pickle.dump([self.target.as_numpy[:],self.target_countN], file)
             else:
@@ -95,7 +98,10 @@ class Tester():
         # Generate test stepper data if it doesn't exist
         test_stepper = test_stepper(op, **stepper_args)
[…]

### 2024-09-23 article_4.tex
Mass Matrix and finite elements

@@
 Again we continue to see that the Lanzcos algorithm outperforms the Arnodli algorithm.
+
+\section{Application to Finite Element Methods}
+Here we discuss the application of the exponential integrator to finite element methods.
+
+We begin with a finite element problem in the weak form:
+\begin{align*}
+\int \frac d{dt}u_j v_j - R(u)_j v_j dx &= \mathcal{L}(u,v)\\
+\text{substituting } u(t,x) &= \sum_{i=1}^n u_i(t)v_i(x)\\
+\text{we get } \int\sum_{i=1}^n(\frac d{dt}u_i(t)v_i(x)v_j(x))dx - R_jv_j(x) &= \sum_{i=1}^n u_i(t)\mathcal{L}(v_i(x)v_j(x))\\
+\frac d{dt}u_i(t)\sum_{i=1}^n\int(v_i(x)v_j(x))dx - R_jv_j(x) &= \sum_{i=1}^n u_i(t)\mathcal{L}(v_i(x)v_j(x))\\
+M\frac d{dt}u(t) - R &= Lu\\
+\frac d{dt}u(t) &= M^{-1}(Lu + R(u))
+\end{align*}
+A clear limiting factor when computing $M^{-1}(Lu + R(u))$ is taking the inverse of $M$, the matrix.
+One way around this is to perform mass lumping on the matrix, where we sum up the values along each row and place them along the diagonal.
+From here computing the inverse is trivial.
+
 \clearpage

### 2024-09-24 parabolicTest.py
Changed Parabolic Tests

@@ -35,7 +35,7 @@ def paraTest1(gridView):
 
 def paraTest2(gridView):
     tauFE = 7e-5 # check if this is correct
-    freq = 6*pi
+    freq = 2*pi
     xExact = as_vector([ cos(freq*x[0]) ])
     exact = lambda t: cos(freq**2*t) * xExact
     dtExact = lambda t: -freq**2*sin(freq**2*t) * xExact

### 2024-09-27 FETester.py
FETester Updated

@@ -131,7 +131,7 @@ class Tester():
         self.run_test(tau, stepper, stepper_args, end_time)
 
 if __name__ == "__main__":
-    if True:
+    if False:
         from allenCahn import dimR, time, sourceTime, domain
         from allenCahn import test2 as problem
         problemName = "Allen Cahn Test2"
@@ -140,25 +140,21 @@ if __name__ == "__main__":
         tau0 = 1e-1
     else:
         from parabolicTest import dimR, time, sourceTime, domain
-        from parabolicTest import paraTest1 as problem
-        problemName = "Parabolic Test1"
-        tau0 = 1e-4  # 2e-3,N:32,Tau:0.002: compare m=5->m=10
-        start_time = 0.0005
-        end_time = 0.001
+        from parabolicTest import paraTest2 as problem
+        problemName = "Parabolic Test2"
+        tau0 = 2e-4  # 2e-3,N:32,Tau:0.002: compare m=5->m=10
+        start_time = 0.00
+        end_time = 0.02
     exp_methods = ["EXPARN", "EXPLAN", "EXPKIOPS", "BE"]
 
     results = []
 
-
-    # exp_methods = ["EXPLAN", "EXPARN", "EXPKIOPS"]
-
-
     domain = list(domain)
     for exp_method in exp_methods:
         print("EXP method:{0}".format(exp_method))
-        for N in [10, 30, 60, 120]:
+        for N in [10, 30, 60]:
             print("N:{0}".format(N))
-            for tau in tau0*np.array([1/2**i for i in range(0, 4)]):
+            for tau in tau0*np.array([1/2**i for i in range(0, 5)]):
                 print("Tau:{0}".format(tau))
 
                 domain[2] = [N,N]
@@ -229,12 +225,12 @@ if __name__ == "__main__":
     markers = ["1", "2", "3", "4"]
     for grid_size in results["Grid Size"].unique():
         if "BE" not in exp_methods:
-            plt.scatter(results["Target N Count"], results["Error L2"], marker="x", label="BE Method")
+            plt.plot(results["Target N Count"], results["Error H1"], marker="x", label="BE Method")
         for i, exp_method in enumerate(exp_methods):
             trimmed_data = results[results["Method"] == exp_method]
             trimmed_data = trimmed_data[trimmed_data["Grid Size"] == grid_size]
 
[…]

### 2024-09-27 article_4.tex
Numerical results for parabolic problem and matrix free methods

@@
 \section {Introduction}
+\subsection{Motivation}
 %Statement of ODE
@@
 This arrives from discretising the PDE in space yielding an ODE with respect to time, with the form shown in \eqref{ODE}.\\
+\subsection{Solvers}
 %ODE Solvers
@@
 \begin{align*}
-u(t+\tau) &\approx u(t) + \tau\dot u(t) = u(t) + \tau(Au(t) + N(u(t)))
+u(t+\tau) &\approx u(t) + \tau\dot u(t) = u(t) + \tau(Au(t) + R(u(t)))
 \end{align*}
@@
 %Computing matrix exponential for A sparse
+\subsection{Computing the Matrix Exponential}
 An important requirement when using this method is to compute approximations of $e^{A}v$ efficiently and accurately for the given sparse matrix $A$ and vector $v$.
@@
 As the matrix exponential will be computed in the lower dimension Kyrlov subspace the computation cost of this step will not be significant.
+\subsection{Matrix Free Methods}
+Another way of reducing computation time is the use of matrix-free methods.
+Frequently to arrive at $\dot u(t) = Au(t) + R(u)$ we will have begun with $\dot u(t) = N(u)$ and linearise around $u(t)$ giving:
+\begin{align*}
+Au &= DN(u)
+&\approx \frac{N(u+\epsilon)-N(u)}{\epsilon}
+\end{align*}
+As a result, we can approximate the action of the matrix  $A$ on $u$ without having to compute $A$ directly or perform a matrix-vector multiplication
+
 \section{Krylov Subspace Methods}
@@
 	  \label{fig:MEe7}
+    \end{minipage}\hfill
+    \begin{minipage}{0.5\textwidth}
+       \centering
+	  \includegraphics[width=\linewidth]{Plots/M v E Results for N=524288.png}
+	  \label{fig:MEe6}
+    \end{minipage}
+\end{figure}
+\begin{figure}[H]
+    \centering
+    \begin{minipage}{0.5\textwidth}
        \centering
@@
 	  \label{fig:MEe7}
+    \end{minipage}\hfill
+    \begin{minipage}{0.5\textwidth}
+       \centering
+	  \includegraphics[width=\linewidth]{Plots/M v Comp Time Results for N=524288.png}
+	  \label{fig:MEe6}
+    \end{minipage}
+\end{figure}
+\begin{figure}[H]
+    \centering
+    \begin{minipage}{0.5\textwidth}
        \centering
@@
     \begin{minipage}{0.5\textwidth}
-       \centering
-	  \includegraphics[width=\linewidth]{Plots/M v E Results for N=524288.png}
-	  \label{fig:MEe6}
-       \centering
-	  \includegraphics[width=\linewidth]{Plots/M v Comp Time Results for N=524288.png}
-	  \label{fig:MEe6}
        \centering
@@
 
-We begin with a finite element problem in the weak form:
+We begin with a finite element problem in the weak form with $V_h$ a finite-dimensional space.
 \begin{align*}
-\int \frac d{dt}u_j v_j - R(u)_j v_j dx &= \mathcal{L}(u,v)\\
-\text{substituting } u(t,x) &= \sum_{i=1}^n u_i(t)v_i(x)\\
-\text{we get } \int\sum_{i=1}^n(\frac d{dt}u_i(t)v_i(x)v_j(x))dx - R_jv_j(x) &= \sum_{i=1}^n u_i(t)\mathcal{L}(v_i(x)v_j(x))\\
-\frac d{dt}u_i(t)\sum_{i=1}^n\int(v_i(x)v_j(x))dx - R_jv_j(x) &= \sum_{i=1}^n u_i(t)\mathcal{L}(v_i(x)v_j(x))\\
+\int \frac d{dt}u(t,x) v_j - R(u) v_j dx &= \mathcal{L}(u,v)\text{Where $v_j$ is a basis of $V_h$}\\
+\text{substituting in } u(t,x) &= \sum_{i=1}^n u_i(t)v_i(x)\\
+\text{we get } \int\sum_{i=1}^n(\frac d{dt}u_i(t)v_i(x)v_j(x))dx - R(u)v_j(x)dx &= \sum_{i=1}^n u_i(t)\mathcal{L}(v_i(x)v_j(x))\\
+\frac d{dt}u_i(t)\sum_{i=1}^n\int v_i(x)v_j(x)dx - \int R(u)v_j(x)dx &= \sum_{i=1}^n u_i(t)\mathcal{L}(v_i(x)v_j(x))\\
+\text{setting } S_{ij} &= \mathcal{L}(v_i(x)v_j(x))\\
+R_j &= \int R(u)v_j(x)dx\\
+M &= \int v_i(x)v_j(x)dx\\
+\text{We get}\\
 M\frac d{dt}u(t) - R &= Lu\\
@@
 
+\subsection{Numerical Results}
+We will now look at the application of the exponential integrator to numerical problems.
+
+We ran our tests over several $N\times N$ grids for various values of $N$ and a selection of values for our time step $\tau$.
+There is an initial "seeding" phase where we use a predetermined stepper to compute an approximation from time $t=0$ to a predetermined seed time $t_s$ and with a seeding time step $\tau_s$ producing an approximation at the seed time of $u_s$.
+From here we can test our different steppers beginning at time $t_s$ with initial condition $u_s$ using a time step $\tau$ and run until an end time $t_e$.
+
+\subsubsection{Parabolic Problem}
+Here we look at the results for a Parabolic Problem given below:
+
+The exact solution is given below:
+\begin{align*}
+\hat u(t,x) &= cos(4\pi t)cos(2\pi x_0)
+\end{align*}
+
+We set $t_s = 0$ and $t_e = 0.02$ and test with $\tau = 2\times 10^{-4}, 1\times 10^{-4}, 5\times 10^{-5}, 2.5\times 10^{-5}, 1.25\times 10^{-5}$.
+For grid sizes, we use $N\times N$ with $N = 10,30,60$.
+
+The metrics we will be comparing are the L2 error given by $\int |u_e - \hat u|$ and the number of calls to the operator $N$ (as we are using a matrix-free method).
+We present the results for the Arnoldi, Lanczos, Kipops\cite{Gaudreault2018} and Backward Euler methods.
+
+\begin{figure}[H]
+	  \includegraphics[width=\linewidth]{FEMethodPlots/Operator Calls V Error for Parabolic Test2 with grid size 100.png}
+\end{figure}
+\begin{figure}[H]
+	  \includegraphics[width=\linewidth]{FEMethodPlots/Operator Calls V Error for Parabolic Test2 with grid size 900.png}
+\end{figure}
+\begin{figure}[H]
+	  \includegraphics[width=\linewidth]{FEMethodPlots/Operator Calls V Error for Parabolic Test2 with grid size 3600.png}
+\end{figure}
+
+From this we see that for $1.25\times 10^{3}$ calls to the
+
+\subsubsection{Allen Cahn}
 \clearpage

### 2024-10-11 article_4.tex
Updated Article

@@
 Various methods have been proposed for this such as by Al-Mohy, Awad H.
-and Higham, Nicholas J\cite{AlMohy2011} which is currently used in the Scipy package as \verb|scipy.sparse.linalg.expm_multiply|.
+and Higham, Nicholas J\cite{AlMohy2011} and Higham, N.
+J., and Al-Mohy, A.
+H.
+\cite{Higham2010} which is currently used in the Scipy package as \verb|scipy.sparse.linalg.expm_multiply|.
 Another class of methods known as Krylov subspace methods have gained traction recently\cite{Moler2003}.
@@
 \begin{align*}
-Au &= DN(u)
+Au &= DN(u)\\
 &\approx \frac{N(u+\epsilon)-N(u)}{\epsilon}
@@
 \subsubsection{Parabolic Problem}
-Here we look at the results for a Parabolic Problem given below:
-
-The exact solution is given below:
+Here we look at the results for a Parabolic Problem given in the weak form below:
 \begin{align*}
-\hat u(t,x) &= cos(4\pi t)cos(2\pi x_0)
+\text{with exact solution }\hat u(t,x) &= cos(4\pi t)cos(2\pi x_0) \text{ we have}\\
+\int \partial_tu v dx&= a(u,v) - l(v)\\
+\text{where }a(u,v) &= \int \nabla u \cdot \nabla v + u \cdot v dx\\
+\text{and }l(v) &= \int\partial_tu - \nabla \cdot \nabla \hat u)  v_0\\
+&+\hat u \cdot vdx\\
+&+\int \nabla \hat u \cdot n v_0 dx
 \end{align*}
@@
 
-From this we see that for $1.25\times 10^{3}$ calls to the
+From this, we see that for about $1.25\times 10^{3}$ calls to the operator, we achieve a lower error than the backward-euler method was able to.
+This corresponds to a tau of $ 5\times 10^{-5}$.
+We also observe that after this value, error for all methods plateaus and even increases.
 
 \subsubsection{Allen Cahn}
+Here we apply the same exponential integrator methods as above to the Allan Cahn equations.
+We give the Allen Cahn equations in weak form below
+
+\begin{align*}
+\int \partial_tu v dx &= a(u,v)\\
+\text {where } a(u,v) &= \int 10^{-4} \nabla u \cdot \nabla v + (u^2-1)uv dx
+\end{align*}
+
+For computing our seed we have $t_s = 4$ and use $\tau = 10^{-2}$ with the Backward Euler method.
+For the true value $\hat u$ we use the Backward Euler method to approximate it with $\tau = 10^{-4}$.
+We set $t_s = 4$ and $t_e = 50$ and test with $\tau = 2.5\times 10^{-2}$.
+For grid sizes, we use $N\times N$ with $N = 60$.
+
+We provide the plot of the seed and at times $t = 20, 50$ for the different methods.
+
+\begin{figure}[H]
+	\centering
+	\includegraphics[width=\linewidth]{FEPics/Allen Cahn Test2_60x60_0.025_Seed BE_EndTime=50.png}
+	\caption{Seed $t=4$}
+\end{figure}
+\begin{figure}[H]
+    \centering
+    \begin{minipage}{0.5\textwidth}
+       \centering
+	  \includegraphics[width=\linewidth]{FEPics/Allen Cahn Test2_60x60_0.025_Target BE_EndTime=20.png}
+	  \caption{True $t=20$}
+    \end{minipage}\hfill
+    \begin{minipage}{0.5\textwidth}
+       \centering
+	  \includegraphics[width=\linewidth]{FEPics/Allen Cahn Test2_60x60_0.025_Target BE_EndTime=50.png}
+	  \caption{True $t=50$}
+    \end{minipage}
+\end{figure}
+\begin{figure}[H]
+    \centering
+    \begin{minipage}{0.5\textwidth}
+       \centering
+	  \includegraphics[width=\linewidth]{FEPics/Allen Cahn Test2_60x60_0.025_BE_EndTime=20.png}
+	  \caption{Backward Euler $t=20$}
+    \end{minipage}\hfill
+    \begin{minipage}{0.5\textwidth}
+       \centering
+	  \includegraphics[width=\linewidth]{FEPics/Allen Cahn Test2_60x60_0.025_BE_EndTime=50.png}
+	  \caption{Backward Euler$t=50$}
+    \end{minipage}
+\end{figure}
+\begin{figure}[H]
+    \centering
+    \begin{minipage}{0.5\textwidth}
+       \centering
+	  \includegraphics[width=\linewidth]{FEPics/Allen Cahn Test2_60x60_0.025_EXPARN_EndTime=20.png}
+	  \caption{Arnoldi $t=20$}
+    \end{minipage}\hfill
+    \begin{minipage}{0.5\textwidth}
+       \centering
+	  \includegraphics[width=\linewidth]{FEPics/Allen Cahn Test2_60x60_0.025_EXPARN_EndTime=50.png}
+	  \caption{Arnoldi $t=50$}
+    \end{minipage}
+\end{figure}
+\begin{figure}[H]
+    \centering
+    \begin{minipage}{0.5\textwidth}
+       \centering
+	  \includegraphics[width=\linewidth]{FEPics/Allen Cahn Test2_60x60_0.025_EXPLAN_EndTime=20.png}
+	  \caption{Lanzcos $t=20$}
+    \end{minipage}\hfill
+    \begin{minipage}{0.5\textwidth}
+       \centering
+	  \includegraphics[width=\linewidth]{FEPics/Allen Cahn Test2_60x60_0.025_EXPLAN_EndTime=50.png}
+	  \caption{Lanzcos $t=50$}
+    \end{minipage}
+\end{figure}
+\begin{figure}[H]
+    \centering
+    \begin{minipage}{0.5\textwidth}
+       \centering
+	  \includegraphics[width=\linewidth]{FEPics/Allen Cahn Test2_60x60_0.025_EXPKIOPS_EndTime=20.png}
+	  \caption{Kiops $t=20$}
+    \end{minipage}\hfill
+    \begin{minipage}{0.5\textwidth}
+       \centering
+	  \includegraphics[width=\linewidth]{FEPics/Allen Cahn Test2_60x60_0.025_EXPKIOPS_EndTime=50.png}
+	  \caption{Kiops $t=50$}
+    \end{minipage}
+\end{figure}
+
+The plots for different methods appear to be largely similar to each other.
+
+\section{Conclusion}
+Kyrlov subspace methods for computing the action of the matrix exponential on the vector appear to outperform some already commonly used methods such as those used in the scipy library regarding computation time.
+As a result, when applied to the numerical solutions of PDEs via finite element methods with the use of exponential integrators, we see that Krylov subspace methods appear to be able to achieve a lower minimum L2 error when applied to parabolic problems than the Backward Euler method, especially for finer grid spacing, implying that they may be more suitable for finner grid spacings.
+We also see that when applied to the Allen Cahn equations we achieve approximately similar results to the Backward Euler method suggesting that they are also useful for this application.
+
+\section{Outlook}
+Further areas requiring investigation include:
+Finding how many dimensions are necessary for the Kyrlov subspace to achieve a good approximation for a given problem.
+Kiops attempts to calculate an optimal size but the Arnoldi and Lanzcos methods achieved comparable results for a lower dimension and were therefore faster.
+
 \clearpage
